---
title: "create_ushist"
author: "Tourism Economics"
date: "Thursday, October 16, 2014"
output: 
  html_document: 
    keep_md: yes
---

This had been pretty similar to create_ushist. To make this, I changed it to import the host Rdata files. I deleted the parts related to opens and closes. And removed stuff about IHG canada and mexico. And commented out the top25 piece. 

Then I went through a more thorough reworking. I realized I could do the calculation of real series across multiple countries.


Setup
```{r readchunk}
#read_chunk('~/Project/R projects/lodfor/scripts/functions.R')
#source('~/Project/R projects/lodfor/scripts/functions_combined.R')
library(arlodr, warn.conflicts=FALSE)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
```


Creates a us historical databank. Combines the STR data with selected macro data and calculates a few series
```{r load}

fpath <- c("~/Project/R projects/lodfor/") 
#macro data
  load(paste(fpath,"output_data/oe_usmac_q.Rdata", sep=""))
# str data
  load(paste(fpath,"output_data/out_str_us_host_q.Rdata", sep="")) 
  load(paste(fpath,"output_data/out_str_us_host_m.Rdata", sep=""))

# put quarterly into a tidy format
temp_str_q <- out_str_us_host_q %>%
  data.frame(date=time(.), .) %>%
  # creates column called segvar that contains the column names, and one next to 
  # it with the values, dropping the time column
  gather(segvar, value, -date, na.rm = FALSE) %>%
  # separates out the variables
  separate(segvar, c("seg", "variable"), sep = "_", extra="merge") %>%
  separate(variable, c("area_sh", "variable"), sep = "_", extra="merge") %>%
  separate(variable, c("country", "variable"), sep = "_", extra="merge") %>%
  # keeps everything else as columns and spreads variable into multiple 
  # columns containing the values
  spread(variable,value)

# put monthly into a tidy format
temp_str_m <- out_str_us_host_m %>%
  data.frame(date=time(.), .) %>%
  # creates column called segvar that contains the column names, and one next to 
  # it with the values, dropping the time column
  gather(segvar, value, -date, na.rm = FALSE) %>%
  # separates out the variables
  separate(segvar, c("seg", "variable"), sep = "_", extra="merge") %>%
  separate(variable, c("area_sh", "variable"), sep = "_", extra="merge") %>%
  separate(variable, c("country", "variable"), sep = "_", extra="merge") %>%
  # keeps everything else as columns and spreads variable into multiple 
  # columns containing the values
  spread(variable,value)
  
```

The initial steps do the quarterly databank. Monthly is done further below.
```{r create_q}
# selects certain series to bring in. Others just stay in macro in case they 
# are needed in future.
temp_usmac <- oe_usmac_q %>%
  data.frame(date=time(oe_usmac_q), oe_usmac_q) %>%
  select(date, 
         us_gdp,
         us_ifix,
         us_cd,
         us_iconstr,
         us_popnipa,
         us_popw,
         us_et,
         us_up,
         us_yhat,
         us_pc,
         us_pedy,
         us_penwall,
         us_cogtp,
         us_cpi,
         us_usrecq,
         can_gdp,
         can_cpi,
         can_pc,
         mex_gdp,
         mex_cpi,
         mex_pc
         ) %>%
  read.zoo %>%
  as.xts
head(temp_usmac)

# merges dataframes. the all.=TRUE piece ensures all the rows
# in the first dataframe are included
#ushist_q <- merge(temp, out_str_us_host_q, all=TRUE) 

######################
#
# indexes everything in ushist_q
#
# ushist_ind_q <- index_q_xts(ushist_q,index_year=2005)
# 
# # look at a graph
# tempa <- ushist_ind_q$totus_occ_sa
# tempb <- ushist_ind_q$upsus_occ_sa
# tempc <- merge(tempa,tempb)
# autoplot(window(tempc, start="2000-01-01", end="2014-10-01"), facets=NULL)

```

##Converts some series to real

First set up the price deflator to use for each country
Index the personal cons price deflator to average 100 in 2014
```{r pcindex}

# US
us_pcindex <- index_q(temp_usmac$us_pc, index_year=2014)
names(us_pcindex) <- "us_pcindex"
autoplot.zoo(us_pcindex)

# Canada
can_pcindex <- index_q(temp_usmac$can_pc, index_year=2014)
names(can_pcindex) <- "can_pcindex"
autoplot.zoo(can_pcindex)

# Mexico
mex_pcindex <- index_q(temp_usmac$mex_pc, index_year=2014)
names(mex_pcindex) <- "mex_pcindex"
autoplot.zoo(mex_pcindex)

# combines as a dataframe
pcindex <- merge(us_pcindex, can_pcindex, mex_pcindex) %>%
  data.frame(date=time(.), .) %>%
  gather(var, value, -date, na.rm = FALSE) %>%
  separate(var, c("country", "variable"), sep = "_", extra="merge") %>%
  mutate(country = ifelse(country == "us", "usa", country)) %>%
  spread(variable, value)

```

Add the price index to the somewhat tidy dataframe and calculate real series
```{r calc_real}

temp_str_q1 <- temp_str_q %>%
  left_join(., pcindex, by=c("date", "country")) %>%
  mutate(adrr = (adr / pcindex)*100) %>%
  mutate(adr_sar = (adr_sa / pcindex)*100) %>%
  mutate(revparr = (revpar / pcindex)*100) %>%
  mutate(revpar_sar = (revpar_sa / pcindex)*100)
  
temp_str_q2 <- temp_str_q1 %>%
  mutate(seg = paste0(seg, area_sh)) %>%
  # drop country because I don't think it is needed anymore
  select(-country, -area_sh) %>%
  gather(variable, value, -seg, -date) %>%
  mutate(segvar = paste(seg, variable, sep="_")) %>%
  select(-seg, -variable) %>%
  spread(segvar, value)

# to remove columns that are all NA and read back to xts
str_q <- temp_str_q2[,colSums(is.na(temp_str_q2))<nrow(temp_str_q2)] %>%
 read.zoo() %>%
 xts()

autoplot.zoo(window(str_q$upachcil_adr_sar, start="2000-01-01", end="2014-10-01"))
autoplot.zoo(window(str_q$upachcil_revpar_sa, start="2000-01-01", end="2014-10-01"))
autoplot.zoo(window(str_q$upachcil_revpar, start="2000-01-01", end="2014-10-01"))
autoplot.zoo(window(str_q$luxusxxx_adr, start="2000-01-01", end="2014-10-01"))

# merges onto ushist_q
ushist_q <- merge(temp_usmac, str_q)

autoplot.zoo(window(ushist_q$upausxxx_adr, start="2000-01-01", end="2014-10-01"))
autoplot.zoo(window(ushist_q$upausxxx_adr_sa, start="2000-01-01", end="2014-10-01"))
autoplot.zoo(window(ushist_q$upausxxx_adr_sar, start="2000-01-01", end="2014-10-01"))

ushist_q_df <- ushist_q %>%
  data.frame(date=time(.), .)


# simple graph
grapht <- ushist_q_df %>%
  select(date, upausxxx_adr, upausxxx_adr_sa, upausxxx_adr_sar) %>%
  read.zoo() %>%
  xts() %>%
  window(.,start="2000-01-01", end="2015-10-01")
autoplot.zoo(grapht, facets = NULL)

# simple graph with variables for selected geo
segment <- c("upanwyny")
var1 <- paste0(segment,"_adr")
var2 <- paste0(segment,"_adr_sa")
var3 <- paste0(segment,"_adr_sar")

colnames(ushist_q)
grapht <- ushist_q_df %>%
  select(date, get(var1), get(var2), get(var3)) %>%
  read.zoo() %>%
  xts() %>%
  window(.,start="2000-01-01", end="2015-10-01")
autoplot.zoo(grapht, facets = NULL)
```


Looking at what's in quarterly databank
```{r look_q}
# which segments or markets are in the data frame, just for observation
# not used anywhere
a <- grep(pattern="_demt", colnames(ushist_q), value=TRUE)
a
a <- gsub(pattern="_demt",replacement="",a)
a

b <- grep(pattern="totusxxx_", colnames(ushist_q), value=TRUE)
b

c <- colnames(ushist_q)
c
```

Create a sum of top 25 metros
```{r top25}
# 
# I commetted this out. my thought is that it may be better to set up a compile_top25 type file to do this
# that can handle seasonally adjusting and putting into real terms

# selectm_list <- c("atl", "bos", "chi", "den", "hou", "los", "mia", "mou",
#                      "nol", "nyc", "org", "orl", "phl", "pho", "sea", "snd",
#                      "snf",  "was")
# 
# selectm_cols <- unique (grep(paste(selectm_list,collapse="|"), 
#                         colnames(ushist_q), value=TRUE))
# 
# # follows the code for creating the annual databank, see comments in next code chunk
# selectm_sum <- data.frame(ushist_q) %>%
#   select(matches("_demt|_supt|_rmrevt")) %>%
#   select(matches(paste(selectm_cols, collapse="|"))) %>%
#   as.xts()
# 
#   # takes the summed data and spreads it into a tidy format with tidyr
#   # creates column called segvar that contains the column names, and one next to 
#   # it with the values, dropping the time column
# selectm_sum <- data.frame(date=time(selectm_sum), selectm_sum)%>% 
#   # creates column called segvar that contains the column names, and one next to 
#   # it with the values, dropping the time column
#   gather(segvar, value, -date, na.rm = FALSE) %>%
#   # in the following the ^ means anything not in the list
#   # with the list being all characters and numbers
#   # so it separates segvar into two colums using sep
#   separate(segvar, c("seggeo", "variable"), sep = "[^[:alnum:]]+") %>%
#   # keeps seg as a column and spreads variable into multiple columns containing
#   # containing the values
#   mutate(seg = substr(seggeo, 1, 3)) %>%
#   mutate(geo = substr(seggeo, 4, 6)) %>%
#   select(-seggeo) %>%
#   spread(variable,value) 
# 
# selectm_sum <- selectm_sum %>%
#   # sets up in dplyr that it will summarize by quarters
#   group_by(date, seg) %>%
#   # calculates top25us as the sum of the markets in the dataframe
#   summarize(slm_demt=sum(demt), slm_supt=sum(supt), 
#             slm_rmrevt=sum(rmrevt)) %>%
#   mutate(slm_occ = slm_demt / slm_supt) %>%
#   mutate(slm_revpar = slm_rmrevt / slm_supt) %>%
#   mutate(slm_adr = slm_rmrevt / slm_demt) %>%
#   # added this for host situation
#   gather(segvar, value, -date, -seg, na.rm = FALSE) %>%
#   mutate(segvar = paste(seg, segvar, sep="")) %>%
#   select(-seg) %>%
#   spread(segvar,value) 
# 
# selectm_sum <- selectm_sum %>%
#   # there's a bit of a bug in dplyr in that the data frame
#   # can't be directly read by read.zoo, something about a
#   # bad entry error that happens. So the short term solution
#   # is to have the as.data.frame step. Here's the bug report
#   # which will presumably be fixed at some point
#   # https://github.com/hadley/dplyr/issues/686
#   as.data.frame() %>%
#   read.zoo() %>%
#   xts()
# 
# ushist_q <- merge(ushist_q, selectm_sum)

```


Create annual databank
```{r create_a}
# start with those that should be summed

# select series that should be converted to annual by summing
# I wrote a regular expression that is looking for certain text strings
# for reference on writing regular expressions, see
# http://www.regular-expressions.info/quickstart.html
suma <- data.frame(ushist_q) %>%
  select(matches("_demt|_supt|_rmrevt")) %>%
  as.xts()

# this function is one I defined, it converts all the columns in 
# an xts object to annual. Must be an xts object to start with
suma <- q_to_a_xts(suma, type="sum")

# takes the summed data and spreads it into a tidy format with
# tidyr and then calculates the occupancy and revpar series
# first needs to go from xts to dataframe
tb2 <- data.frame(date=time(suma), suma)%>% 
  # creates column called segvar that contains the column names, and one next to 
  # it with the values, dropping the time column
  gather(segvar, value, -date, na.rm = FALSE) %>%
  # in the following the ^ means anything not in the list
  # with the list being all characters and numbers
  # so it separates segvar into two colums using sep
  separate(segvar, c("seg", "variable"), sep = "[^[:alnum:]]+") %>%
  # keeps seg as a column and spreads variable into multiple columns containing
  # containint the values
  spread(variable,value) %>%
  # adds new calculated column
  mutate(occ = demt / supt) %>%
  # adds another column
  mutate(revpar = rmrevt / supt) %>%
  mutate(adr = rmrevt / demt)

a <- tb2$segvar
b <- unique(a)
b

# takes it from a tidy format and melts it, and then creates the unique
# variable names and then reads into a zoo object spliting on the 
# second column
#a <- melt(tb2, id=c("date","seg"), na.rm=FALSE)
a <- tb2 %>%
  gather(variable, value, -date, -seg)
a$variable <- paste(a$seg, "_", a$var, sep='')
a$seg <- NULL
ushist_a <- xts(read.zoo(a, split = 2))

# looking at a few graphs
autoplot.zoo(ushist_a$luxusxxx_revpar)
autoplot.zoo(window(ushist_a$totusxxx_occ, start=as.Date("1987-01-01"), end=as.Date("2015-01-01")))
```

Creating monthly historical databank
```{r create_m}
# not that much that needs to be done
temp_str_m1 <- temp_str_m %>%
  mutate(seg = paste0(seg, area_sh)) %>%
  # drop country because I don't think it is needed anymore
  select(-country, -area_sh) %>%
  gather(variable, value, -seg, -date) %>%
  mutate(segvar = paste(seg, variable, sep="_")) %>%
  select(-seg, -variable) %>%
  spread(segvar, value)

# to remove columns that are all NA and read back to xts
ushist_m <- temp_str_m1[,colSums(is.na(temp_str_m1))<nrow(temp_str_m1)] %>%
 read.zoo() %>%
 xts()

```


```{r rename, echo=FALSE}
ushist_host_q <- ushist_q
ushist_host_m <- ushist_m
ushist_host_a <- ushist_a
 
c <- colnames(ushist_host_q)
c
```

Writing outputs
```{r write_output, echo=FALSE}
# quarterly
  # writes csv versions of the output files
  write.zoo(ushist_host_q, file=paste(fpath, "output_data/ushist_host_q.csv", sep=""), sep=",")
  
  # saves Rdata versions of the output files
  save(ushist_host_q, file=paste(fpath, "output_data/ushist_host_q.Rdata", sep=""))
 
# monthly
  # writes csv versions of the output files
  write.zoo(ushist_host_m, file=paste(fpath,"output_data/ushist_host_m.csv",sep=""), sep=",")
  
  # saves Rdata versions of the output files
  save(ushist_host_m, file=paste(fpath,"output_data/ushist_host_m.Rdata", sep=""))

# annual
  # writes csv versions of the output files
  write.zoo(ushist_host_a, file=paste(fpath,"output_data/ushist_host_a.csv",sep=""), sep=",")
  
  # saves Rdata versions of the output files
  save(ushist_host_a, file=paste(fpath,"output_data/ushist_host_a.Rdata", sep=""))
```
