---
title: "Untitled"
author: "Tourism Economics"
date: "November 13, 2015"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
require(directlabels)
require(rmarkdown)
require(knitr)
require(RColorBrewer)
require(gridExtra)
require(gtable)
require(Cairo)

# If I put autodep=TRUE as an argument in the following then I get
# some issues with the graphs, such as grey backgrounds and changes
# in labeling. Maybe it somehow impacts the theme?
#knitr::opts_chunk$set(cache=TRUE, cache.path='output_data/cache/', dpi=96)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, cache.path='../output_data/cache/', autodep=TRUE, fig.path='../output_data/figure_exp_seasonal_short/fig-',  fig.show='asis', fig.keep='high')
# tells it to keep going even if there is an error
knitr::opts_chunk$set(error = TRUE)

```


```{r readchunk, echo=FALSE, message=FALSE, warning=FALSE}
read_chunk('~/Project/R projects/lodfor/scripts/functions.R')
source('~/Project/R projects/lodfor/scripts/functions.R')

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("~/Project/R projects/lodfor/output_data/ushist_m.Rdata")
```

```{r define_theme, echo=FALSE, message=FALSE, warning=FALSE}
#Displays palette and sets the ts1 theme

#display.brewer.pal(10, "RdBu")
mypallete <- brewer.pal( 10 , "RdBu" )
colors_ts1 <- scale_colour_manual(values = mypallete[c(10, 1, 5, 4)])
# I don't use the following, was an example I saw somewhere maybe as a way to 
# interpolate additional palette colors
#mypal <- colorRampPalette( brewer.pal( 10 , "RdBu" ) )
theme_ts1 <- function(base_size = 12, base_family = "") {
  theme_classic(base_size = base_size) %+replace%
    theme(                                 
      panel.grid.major.x=element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.minor.y=element_blank(),
      axis.ticks.y=element_line(size=.2),
      axis.ticks.x=element_line(size=.2),
      panel.background=element_blank(),
      legend.title=element_blank(),
      legend.key=element_rect(fill="white", colour = "white"),
      legend.key.size=unit(1, "cm"),
      legend.text=element_text(size=rel(1)),
      #legend.position = c(1, .1),
      legend.justification = "right",
      axis.line=element_line(size=.2),
      axis.title.y=element_blank(),
      axis.title.x=element_blank(),
      axis.text=element_text(color="black",size=rel(.7)),
      # following seems to adjust y-axis title
      plot.title=element_text(size=base_size * .8, face="plain",hjust=0,vjust=1)
    #theme(plot.title = element_text(lineheight=.8, face="bold"))
    )
}

```

```{r set_theme, echo=FALSE, message=FALSE, warning=FALSE}
theme_set(theme_ts1())
#this is also a useful theme to keep in mind
#theme_classic()
# this function gets the settings used by the current theme
#theme_get()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# shorten dates
ushist_m_1 <- ushist_m["1987-01-01/2015-09-01"]

# sets up a single series to use for a few examples
temp_df <- ushist_m_1 %>%
  data.frame(date=time(.), .) %>%
  select(date, totus_demd)

x <- temp_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")
head(x)
temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# uses default settings
m1 <- seas(temp_ser_ts, 
           forecast.maxlead = 30,# extends 30 quarters ahead
           seats.appendfcst="yes") 
summary(m1)


m2 <- seas(x = temp_ser_ts,
     regression.variables = c("td", "easter[1]", "ls2001.Sep"), 
     arima.model = "(1 1 0)(1 1 1)", 
     regression.aictest = NULL, 
    outlier = NULL, 
    transform.function = "log")
# the result is the same as the default call
summary(m2)

```


## fundamental identities of seasonal adjustment
Y = T * I * (S * TD)
all.equal(AirPassengers,  
series(m, "seats.trend") *
         series(m, "seats.irregular") * series(m, "seats.adjustfac"))


Name   |   Small   | Description of table  
-------|----------|------------------------------
trend  |   s12    | final SEATS trend component  
seasonal |  s10    | final SEATS seasonal component  
irregular | s13   |  final SEATS irregular component  
seasonaladj | s11 | final SEATS seasonal adjustment component   
transitory | s14 | final SEATS transitory component   
adjustfac | s16 | final SEATS combined adjustment factors  
adjustmentratio | s18 | final SEATS adjustment ratio  
trendfcstdecomp | tfd | forecast of the trend component  
seasonalfcstdecomp |  sfd | forecast of the seasonal component  
seriesfcstdecomp | ofd | forecast of the series component  
seasonaladjfcstdecomp | afd | forecast of the final SEATS seasonal adjustment  
transitoryfcstdecomp | yfd | forecast of the transitory component  
seasadjconst |  sec | final SEATS seasonal adjustment with constant term included  
trendconst | stc | final SEATS trend component with constant term included  
totaladjustment | sta |total adjustment factors for SEATS seasonal adjustment  

```{r}

# temp_ser_fullout <- data_frame(
#    date = as.Date(time(temp_ser_ts)),
#    #original = temp_ser_ts,
#    oa1 = original(m1),
#    sa1 = final(m1),
#    sa1_trend = series(m1, "s12"),
#    sa1_seasonal = series(m1, "s10"),
#    sa1_irregular = series(m1, "s13"),
#    sa1_sf = series(m1, "s16")
#    )

# when I append the forecast, I think I end up with series of 
# different lengths so rather than use the approach above, 
# I had to switch to the approach below

  tempdata_y <- as.xts(temp_ser_ts, start = start(x), end = end(x))
  tempdata_oa1 <- as.xts(original(m1))
  tempdata_sa1 <- as.xts(final(m1)) # seasonally adjusted series
  tempdata_sa1_trend <- as.xts(series(m1, "s12"))
  tempdata_sa1_seasonal = as.xts(series(m1, "s10"))
  tempdata_sa1_irregular = as.xts(series(m1, "s13"))
  tempdata_sa1_sf = as.xts(series(m1, "s16"))

  # names the objects
  names(tempdata_y) <- "y"
  names(tempdata_oa1) <- "oa1"
  names(tempdata_sa1) <- "sa1"
  names(tempdata_sa1_trend) <- "sa1_trend" 
  names(tempdata_sa1_seasonal) <- "sa1_seasonal" 
  names(tempdata_sa1_irregular) <- "sa1_irregular"
  names(tempdata_sa1_sf) <- "sa1_sf" 
  
  # merges the adjusted series onto the existing xts object with the unadjusted
  # series
  temp_ser_fullout <- merge(tempdata_y, tempdata_oa1, tempdata_sa1, 
                            tempdata_sa1_trend, tempdata_sa1_seasonal,   
                            tempdata_sa1_irregular, tempdata_sa1_sf) %>%
      data.frame(date=as.Date(time(.)), ., row.names=NULL)

temp_ser_fullout <- temp_ser_fullout %>%
  mutate(oa1_a = sa1_trend * sa1_irregular *sa1_sf) %>%
  mutate(oa1_b = sa1 * sa1_sf) 

```

code | description
-----|----------------------
oa1 | original series  
sa1 | seasonally adjusted series  
sa1_trend | trend component (s12)
sa1_seasonal | seasonal component (s10)  
sa1_irregular | irregular component (s13)  
sa1_sf | final combined (seasonal/trading day/holiday) factors (s16)  
oa1_a | original = trend * irregular * combined (seasonal/trading day/holiday) adjustment factors
oa1_b | original = seasonally adjusted series * combined (seasonal/trading day/holiday) adjustment factors

### Example calculations
In this table we have two ways to get back to the original series.   
* oa1_a = trend * irregular * combined (seasonal/trading day/holiday) adjustment factors or   
* oa1_b = seasonally adjusted series * combined (seasonal/trading day/holiday) adjustment factors
  

```{r}

a <- temp_ser_fullout %>%
  filter(date > as.Date("1987-01-01") & date < as.Date("2015-09-01")) 

kable(tail(a, n=12), digits = 4, row.names = FALSE)



```

## So which are the seasonal factors?

It's sometimes hard to get a clear understanding from the documentation.  
I've seen reference to the "combined" factors in the D16 (and presumably also s16) table as being the __combined__ (seasonal/trading day/holiday) factors. 

There is also reference to a "total adjustment factors for SEATS seasonal adjustment".

```{r}

temp_ser_fullout <- data_frame(
   date = as.Date(time(temp_ser_ts)),
   #original = temp_ser_ts,
   oa1 = original(m1),
   sa1 = final(m1),
   sa1_trend = series(m1, "s12"),
   sa1_seasonal = series(m1, "s10"),
   sa1_irregular = series(m1, "s13"),
   sa1_sf = series(m1, "s16"),
   sa1_total = series(m1, "sta")
   )

kable(tail(temp_ser_fullout, n=12), digits = 4)

```



```{r}

# comparing two seasonal models

temp_sersa <- data_frame(
   date = as.Date(time(temp_ser_ts)),
   #original = temp_ser_ts,
   oa1 = original(m1),
   sa1 = final(m1),
   oa2 = original(m2),
   sa2 = final(m2)
   )
```


### graph title

```{r}


temp <- temp_sersa %>%
  gather("variable", "value", -date) %>% 
  mutate(variable = gsub("SA", "Seasonally adjusted by SEATS", variable)) %>%
  filter(date > "2012-01-01") %>%
  filter(variable %in% c("oa1", "sa1"))

p1 <- temp %>%
   ggplot()
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40"))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + scale_y_continuous("axis title")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1


```

