---
title: "Seasonal adjustment Intro"
author: "Tourism Economics"
date: "January 9, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
# require(directlabels)
# require(rmarkdown)
# require(gridExtra)
# require(gtable)
# require(Cairo)

library(arlodr)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
#Sys.setenv(X13_PATH = "C:/Aran Installed/x13ashtml")
library(seasonal, warn.conflicts=FALSE)
require(xtable)
require(ggplot2)
require(knitr)
require(RColorBrewer)
require(scales)

# If I put autodep=TRUE as an argument in the following then I get
# some issues with the graphs, such as grey backgrounds and changes
# in labeling. Maybe it somehow impacts the theme?
#knitr::opts_chunk$set(cache=TRUE, cache.path='output_data/cache/', dpi=96)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, cache.path='../output_data/cache/', autodep=TRUE, fig.path='../output_data/figure_exp_seasonal_short/fig-',  fig.show='asis', fig.keep='high')
# tells it to keep going even if there is an error
knitr::opts_chunk$set(error = TRUE)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("~/Project/R projects/lodfor/output_data/ushist_m.Rdata")
load("~/Project/R projects/lodfor/output_data/ushist_q.Rdata")

```

```{r define_theme, echo=FALSE, message=FALSE, warning=FALSE}
#Displays palette and sets the ts1 theme

#display.brewer.pal(10, "RdBu")
mypallete <- brewer.pal( 10 , "RdBu" )
colors_ts1 <- scale_colour_manual(values = mypallete[c(10, 1, 5, 4)])
# I don't use the following, was an example I saw somewhere maybe as a way to 
# interpolate additional palette colors
#mypal <- colorRampPalette( brewer.pal( 10 , "RdBu" ) )
theme_ts1 <- function(base_size = 12, base_family = "") {
  theme_classic(base_size = base_size) %+replace%
    theme(                                 
      panel.grid.major.x=element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.minor.y=element_blank(),
      axis.ticks.y=element_line(size=.2),
      axis.ticks.x=element_line(size=.2),
      panel.background=element_blank(),
      legend.title=element_blank(),
      legend.key=element_rect(fill="white", colour = "white"),
      legend.key.size=unit(1, "cm"),
      legend.text=element_text(size=rel(1)),
      #legend.position = c(1, .1),
      legend.justification = "right",
      axis.line=element_line(size=.2),
      axis.title.y=element_blank(),
      axis.title.x=element_blank(),
      axis.text=element_text(color="black",size=rel(.7)),
      # following seems to adjust y-axis title
      plot.title=element_text(size=base_size * .8, face="plain",hjust=0,vjust=1)
    #theme(plot.title = element_text(lineheight=.8, face="bold"))
    )
}
theme_set(theme_ts1())
```



```{r echo=FALSE, message=FALSE, warning=FALSE}

# shorten dates
ushist_m_1 <- ushist_m["1987-01-01/2015-09-01"]

# sets up a single series to use for a few examples
demo_df <- ushist_m_1 %>%
  data.frame(date=time(.), .) %>%
  select(date, totus_demd)

```


### Average daily demand, monthly observations (2000 to 2015)  
```{r}
temp <- demo_df %>%
 # data.frame(date=as.Date(time(.)), .) %>%
  gather("variable", "value", -date) %>%
  filter(date >= "2000-01-01" & date <= "2015-09-01") %>%
  mutate(month = lubridate::month(date, label = TRUE)) 
  
p1 <- temp %>%
   ggplot()
p1 <- p1 + geom_point(data = temp, aes(x = date, y=value, shape=variable), size=1.5)
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value), size = .7) 
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + ggtitle("Demand, average daily")
p1

```


### Average daily demand, monthly observations (2012 to 2015)  
```{r echo=FALSE, message=FALSE, warning=FALSE}

temp <- demo_df %>%
 # data.frame(date=as.Date(time(.)), .) %>%
  gather("variable", "value", -date) %>%
  filter(date >= "2012-01-01" & date <= "2015-09-01") %>%
  mutate(month = lubridate::month(date, label = TRUE)) 
  
p1 <- temp %>%
   ggplot()
p1 <- p1 + geom_point(data = temp, aes(x = date, y=value, shape=variable), size=1.5)
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value), size = .7) 
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + ggtitle("Demand, average daily")
p1

```


### Percentage change from prior year, daily demand by month (2012 to 2015)
Last observation is September. 
The timing of Labor Day negatively impacted September 2014 and then positively impacted September 2015 (Labor Day was Sep. 7, 2015, compared to Sep. 1, 2014).
```{r echo=FALSE, message=FALSE, warning=FALSE}

temp <- demo_df %>%
 # data.frame(date=as.Date(time(.)), .) %>%
  gather("variable", "value", -date) %>%
  filter(date >= "2011-01-01" & date <= "2015-09-01") %>%
  mutate(pchya = value/lag(value,n=12) - 1) %>%
  filter(date >= "2012-01-01" & date <= "2015-09-01")
  
p1 <- temp %>%
   ggplot()
p1 <- p1 + geom_point(data = temp, aes(x = date, y=pchya, shape=variable), size=1.5)
p1 <- p1 + geom_line(data = temp, aes(x = date, y = pchya), size = .7) 
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + ggtitle("Demand, average daily, percentage change from prior year")
p1 <- p1 + scale_y_continuous(labels=percent)
p1 <- p1 + scale_x_date(date_breaks = "3 month", date_minor_breaks = "1 month", date_labels = "%b %y")
p1

```


### Average daily demand by month (2012 to 2015)
```{r}
temp <- demo_df %>%
 # data.frame(date=as.Date(time(.)), .) %>%
  gather("variable", "value", -date) %>%
  filter(date >= "2012-01-01" & date <= "2015-09-01") %>%
  mutate(month = lubridate::month(date, label = TRUE)) 

p1 <- temp %>%
   ggplot()
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value), size = .7) 
p1 <- p1 + geom_point(data = temp, aes(x = date, y=value, shape=variable), size=1.5)
p1 <- p1 + facet_wrap(~month, nrow = 1)
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40"))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Demand, average daily")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_x_date(date_breaks = "2 year", labels = date_format("%y"))
p1
```


### Average daily demand by month (2000 to 2015)
```{r}
temp_ts <- demo_df %>%
  filter(date >= "2000-01-01" & date <= "2015-09-01") %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon") %>%
  as.ts(., start = start(.))

monthplot(temp_ts, main="Demand by month", ylab="Demand, average daily")

# works but not needed
# temp_a <- temp %>%
#   group_by(month) %>%
#  summarise(mean=mean(value))
# p1 <- temp_a %>%
#    ggplot()
# p1 <- p1 + geom_point(data = temp_a, aes(x = month, y = mean), size = 4) 
# p1
```


## Prepare a seasonally adjusted series  

```{r echo=FALSE, message=FALSE, warning=FALSE}

temp_ser_ts <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon") %>%
  as.ts(., start = start(.), end = end(.))

# uses default settings
m1 <- seas(temp_ser_ts)
#summary(m1)
#plot(m1)
#plot(final(m1))

## this works, but ends up with ts objects in dataframe
# temp_sersa <- data_frame(
#    date = as.Date(time(temp_ser_ts)),
#    Original = temp_ser_ts,
#    SA = final(m1)) 

##
# this also works, but is a bit complicated, 
# create dataframe of original and sa series
# temp_ser <- temp_ser_ts %>%
#   as.xts() %>%
#   data.frame(date=time(.), .) %>%
#   mutate(date = as.Date(date))
# colnames(temp_ser) <- c("date", "original")
# temp_final <- (final(m1)) %>%
#   as.xts() %>%
#   data.frame(date=time(.), .) %>%
#   mutate(date = as.Date(date))
# colnames(temp_final) <- c("date", "sa")
# temp_sersa <- full_join(temp_ser, temp_final, by="date")

##
# this works, and is a bit more clear
  # tempdata_oa <- temp_ser_ts %>% as.xts(., start = start(.), end = end(.))
  # tempdata_sa <- as.xts(final(m1)) # seasonally adjusted series
  # # names the objects
  # names(tempdata_oa) <- "original"
  # names(tempdata_sa) <- "sa"
  # 
  # # merges the adjusted series onto the existing xts object with the unadjusted
  # # series
  # temp_sersa <- merge(tempdata_oa, tempdata_sa) %>%
  #     data.frame(date=as.Date(time(.)), ., row.names=NULL)
  
##
# this works, and is the most clear
temp_sersa <- merge(
   # creates and merges invididual zoo series
   original = as.zoo(original(m1)),
   sa = as.zoo(final(m1)),
   sa_sf = as.zoo(series(m1, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.))) %>%
   data.frame(date=as.Date(time(.)), ., row.names=NULL)

```


### Original and seasonally adjusted series (1987 to 2015)
```{r echo=FALSE, message=FALSE, warning=FALSE}

temp <- temp_sersa %>%
   gather("variable", "value", -date) %>% 
   filter(variable != "sa_sf") %>%
   mutate(variable = gsub("sa", "Seasonally adjusted", variable)) 

p1 <- temp %>%
   ggplot() 
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_color_manual(values = c("grey60", mypallete[10:10])) 
p1

```


### Original and seasonally adjusted series (2012 to 2015)
```{r}
temp <- temp_sersa %>%
  gather("variable", "value", -date) %>% 
  filter(variable != "sa_sf") %>%
  mutate(variable = gsub("sa", "Seasonally adjusted", variable)) %>%
  mutate(variable = gsub("original", "Original", variable)) %>%
  filter(date >= as.Date("2012-01-01"))

p1 <- temp %>%
   ggplot() 
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_color_manual(values = c("grey60", mypallete[10:10])) 
p1

```


### Seasonally adjusted series (2012 to 2015)
```{r}

temp <- temp_sersa %>%
  gather("variable", "value", -date) %>% 
  filter(variable != "sa_sf") %>% 
  filter(variable == "sa") %>%
  mutate(variable = gsub("sa", "Seasonally adjusted", variable)) %>%
  filter(date >= as.Date("2012-01-01")) 

p1 <- temp %>%
   ggplot() 
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value, color=variable), size = 1) 
p1 <- p1 + geom_point(data = temp, aes(x = date, y=value, shape=variable), size=1.5, show.legend=FALSE)
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Demand, average daily")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_color_manual(values = c("grey60", mypallete[10:10])) 
p1


```


```{r}
#works but not needed

# temp <- temp_sersa %>%
#   mutate(OA_pchya = original/lag(original,n=12) - 1) %>%
#   gather("variable", "value", -date) %>% 
#   mutate(variable = gsub("sa", "Seasonally adjusted", variable)) %>%
#   filter(date >= as.Date("2012-01-01")) %>%
#   mutate(variable = gsub("OA_pchya", "Original, yoy percentage change", variable)) %>%
#   filter(variable != "original")
# 
# p1 <- temp %>%
#    ggplot() 
# p1 <- p1 + geom_line(data = temp, aes(x = date, y = value, color=variable), size = 1) 
# p1 <- p1 + facet_wrap( ~ variable, ncol=1,scales="free_y")
# p1 <- p1 + geom_point(data = temp, aes(x = date, y=value, shape=variable), size=1.5, show.legend=FALSE)
# p1 <- p1 + labs(colour = "", x = "")
# p1 <- p1 + ggtitle("Demand, average daily")
# p1 <- p1 + theme(legend.position = "bottom")
# p1 <- p1 + scale_color_manual(values = c("grey60", mypallete[10:10]))
# p1

```

### Percentage change from prior year (2007 to 2015)
```{r}
temp <- temp_sersa %>%
  mutate(OA_pchya = original/lag(original,n=12) - 1) %>%
  gather("variable", "value", -date) %>% 
  filter(variable != "sa_sf") %>%
  mutate(variable = gsub("sa", "Seasonally adjusted", variable)) %>%
  filter(date >= as.Date("2007-01-01")) %>%
  mutate(variable = gsub("OA_pchya", "Original, yoy percentage change", variable)) %>%
  filter(variable != "original") %>%
  filter(variable == "Original, yoy percentage change")

p1 <- temp %>%
   ggplot() 
p1 <- p1 + geom_point(data = temp, aes(x = date, y=value, shape=variable), size=1.5, show.legend=FALSE)
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Demand, average daily")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_y_continuous(labels=percent)
p1 <- p1 + scale_color_manual(values = c(mypallete[10:10]))
p1 <- p1 + scale_x_date(date_breaks = "1 year", date_labels = "%Y")
p1

```


### Seasonally adjusted  
```{r}
temp <- temp_sersa %>%
  mutate(OA_pchya = original/lag(original,n=12) - 1) %>%
  gather("variable", "value", -date) %>% 
  filter(variable != "sa_sf") %>%
  mutate(variable = gsub("sa", "Seasonally adjusted", variable)) %>%
  filter(date >= as.Date("2007-01-01")) %>%
  mutate(variable = gsub("OA_pchya", "Original, yoy percentage change", variable)) %>%
  filter(variable != "original") %>%
  filter(variable == "Seasonally adjusted")

p1 <- temp %>%
   ggplot() 
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value, color=variable), size = 1) 
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Demand, average daily")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_color_manual(values = c(mypallete[10:10]))
p1 <- p1 + scale_x_date(date_breaks = "1 year", date_labels = "%Y")
p1
```


### Seasonally adjusted, with smoothed series
```{r}
temp <- temp_sersa %>%
  mutate(OA_pchya = original/lag(original,n=12) - 1) %>%
  gather("variable", "value", -date) %>% 
  filter(variable != "sa_sf") %>%
  mutate(variable = gsub("sa", "Seasonally adjusted", variable)) %>%
  filter(date >= as.Date("2005-01-01")) %>%
  mutate(variable = gsub("OA_pchya", "Original, yoy percentage change", variable)) %>%
  filter(variable != "original") %>%
  filter(variable == "Seasonally adjusted")

start_xaxis <- as.Date("2007-01-01") # for x-axis
end_xaxis <- as.Date("2015-09-01") # for x-axis

p1 <- temp %>%
   ggplot() 
p1 <- p1 + geom_line(data = temp, aes(x = date, y = value, color=variable), size = 1)
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Demand, average daily")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_color_manual(values = c("grey60"))
p1 <- p1 + scale_x_date(date_breaks = "1 year", date_labels = "%Y")
p1 <- p1 + stat_smooth(data = temp, aes(x = date, y = value, color=variable), 
               color=mypallete[10:10], span=.2, se = FALSE)
p1 <- p1 + coord_cartesian(xlim = c(start_xaxis, end_xaxis), ylim = NULL)
p1

```

