---
title: "Seasonal adjustment demo"
author: "Tourism Economics"
date: "January 9, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
# require(directlabels)
# require(rmarkdown)
# require(gridExtra)
# require(gtable)
# require(Cairo)

library(arlodr)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
#Sys.setenv(X13_PATH = "C:/Aran Installed/x13ashtml")
library(seasonal, warn.conflicts=FALSE)
require(xtable)
require(ggplot2)
require(knitr)
require(RColorBrewer)
require(scales)

# If I put autodep=TRUE as an argument in the following then I get
# some issues with the graphs, such as grey backgrounds and changes
# in labeling. Maybe it somehow impacts the theme?
#knitr::opts_chunk$set(cache=TRUE, cache.path='output_data/cache/', dpi=96)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, cache.path='../output_data/cache/', autodep=TRUE, fig.path='../output_data/figure_exp_seasonal_short/fig-',  fig.show='asis', fig.keep='high')
# tells it to keep going even if there is an error
knitr::opts_chunk$set(error = TRUE)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("~/Project/R projects/lodfor/output_data/ushist_m.Rdata")
load("~/Project/R projects/lodfor/output_data/ushist_q.Rdata")

```

```{r define_theme, echo=FALSE, message=FALSE, warning=FALSE}
#Displays palette and sets the ts1 theme

#display.brewer.pal(10, "RdBu")
mypallete <- brewer.pal( 10 , "RdBu" )
colors_ts1 <- scale_colour_manual(values = mypallete[c(10, 1, 5, 4)])
# I don't use the following, was an example I saw somewhere maybe as a way to 
# interpolate additional palette colors
#mypal <- colorRampPalette( brewer.pal( 10 , "RdBu" ) )
theme_ts1 <- function(base_size = 12, base_family = "") {
  theme_classic(base_size = base_size) %+replace%
    theme(                                 
      panel.grid.major.x=element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.minor.y=element_blank(),
      axis.ticks.y=element_line(size=.2),
      axis.ticks.x=element_line(size=.2),
      panel.background=element_blank(),
      legend.title=element_blank(),
      legend.key=element_rect(fill="white", colour = "white"),
      legend.key.size=unit(1, "cm"),
      legend.text=element_text(size=rel(1)),
      #legend.position = c(1, .1),
      legend.justification = "right",
      axis.line=element_line(size=.2),
      axis.title.y=element_blank(),
      axis.title.x=element_blank(),
      axis.text=element_text(color="black",size=rel(.7)),
      # following seems to adjust y-axis title
      plot.title=element_text(size=base_size * .8, face="plain",hjust=0,vjust=1)
    #theme(plot.title = element_text(lineheight=.8, face="bold"))
    )
}
theme_set(theme_ts1())
```

```{r}

segrun <- c("totus_demd")

```


### Set up data to run using: `r segrun`
```{r echo=FALSE, message=FALSE, warning=FALSE}

# shorten dates
ushist_m_1 <- ushist_m["1987-01-01/2015-09-01"]

# sets up a single series to use for a few examples
demo_df <- ushist_m_1 %>%
  data.frame(date=time(.), .) %>%
  select_("date", segrun)

```
  
## Run seasonal adjustment
### Summary results  
Uses default settings. For example, includes a term for Easter.
```{r}
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# uses default settings
m1 <- seas(temp_ser_ts, 
           forecast.maxlead = 30,# extends 30 quarters ahead
           seats.appendfcst="yes") 
summary(m1)

```

```{r}
# create example dataframe with various series
  tempdata_y <- as.xts(temp_ser_ts, start = start(x), end = end(x))
  tempdata_oa1 <- as.xts(original(m1))
  tempdata_sa1 <- as.xts(final(m1)) # seasonally adjusted series
  tempdata_sa1_trend <- as.xts(series(m1, "s12"))
  tempdata_sa1_seasonal = as.xts(series(m1, "s10"))
  tempdata_sa1_irregular = as.xts(series(m1, "s13"))
  tempdata_sa1_sf = as.xts(series(m1, "s16"))

  # names the objects
  names(tempdata_y) <- "y"
  names(tempdata_oa1) <- "oa1"
  names(tempdata_sa1) <- "sa1"
  names(tempdata_sa1_trend) <- "sa1_trend" 
  names(tempdata_sa1_seasonal) <- "sa1_seasonal" 
  names(tempdata_sa1_irregular) <- "sa1_irregular"
  names(tempdata_sa1_sf) <- "sa1_sf" 
  
  # merges the adjusted series onto the existing xts object with the unadjusted
  # series
  temp_ser_fullout_a <- merge(tempdata_y, tempdata_oa1, tempdata_sa1, 
                            tempdata_sa1_trend, tempdata_sa1_seasonal,   
                            tempdata_sa1_irregular, tempdata_sa1_sf) %>%
      data.frame(date=as.Date(time(.)), ., row.names=NULL)

temp_ser_fullout <- temp_ser_fullout_a %>%
  mutate(oa1_a = sa1_trend * sa1_irregular *sa1_sf) %>%
  mutate(oa1_b = sa1 * sa1_sf) 
```


## Series codes for use in examples

code | description
-----|----------------------
oa1 | original series  
sa1 | seasonally adjusted series  
sa1_sf | final combined (seasonal/trading day/holiday) factors (s16)  
oa1_b | original = seasonally adjusted series \* combined (seasonal/trading day/holiday) adjustment factors

### Example calculations
In this table we can get back to the original series as follows:  

- oa1_b = sa1 \* sa1_sf  
  
which is equivalent to:  
  
- oa1_b = seasonally adjusted series \* combined (seasonal/trading day/holiday) adjustment factors
  
  
```{r}

# full table
# a <- temp_ser_fullout %>%
#   filter(date > as.Date("1987-01-01") & date < as.Date("2015-09-01"))

# simple table  
a <- temp_ser_fullout %>%
  filter(date > as.Date("1987-01-01") & date < as.Date("2015-09-01")) %>%
  select(-y, -sa1_trend, -sa1_seasonal, -sa1_irregular, -oa1_a)
  
kable(tail(a, n=12), digits = 4, row.names = FALSE)

```




### Seasonal factors for September
Take the estimate of the annualized level, and multiply by the seasonal factor, to get the actual level
```{r}

# simple table, show september
temp <- temp_ser_fullout %>%
  filter(date >= as.Date("1987-01-01") & date <= as.Date("2015-09-01")) %>%
  mutate(month = lubridate::month(date, label = TRUE, abbr = TRUE)) %>%
  filter(month == "Sep") %>%
  select(-y, -sa1_trend, -sa1_seasonal, -sa1_irregular, -oa1_a)
  
kable(tail(temp, n=12), digits = 4, row.names = FALSE)

```


### Seasonal factors by month (1987 to 2015)
```{r}

temp <- temp_ser_fullout %>%
  filter(date >= as.Date("1987-01-01") & date <= as.Date("2015-09-01")) %>%
  select(date, sa1_sf) %>%
    gather("variable", "value", -date) %>%
    mutate(month = lubridate::month(date, label = TRUE, abbr = TRUE))

p1 <- temp %>%
   ggplot()
#p1 <- p1 + geom_line(data = temp, aes(x = date, y = value), size = .7) 
p1 <- p1 + geom_point(data = temp, aes(x = date, y=value, shape=variable), size=1)
p1 <- p1 + facet_wrap(~month, nrow = 1)
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40"))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Seasonal factors")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_x_date(date_breaks = "10 year", labels = date_format("%y"))
p1
```


# How does it help us?  

## Converting quarterly to monthly using seasonal factors


```{r}

# getting data ready

# shorten dates
ushist_m_1 <- ushist_m["1987-01-01/2015-09-01"]
ushist_q_1 <- ushist_q["1987-01-01/2015-09-01"]

# sets up a single series to use for a few examples
x <- ushist_m_1 %>%
  data.frame(date=time(.), .) %>%
  select_("date", segrun) %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")
tempm_ts <- as.ts(x, start = start(x), end = end(x))

x <- ushist_q_1 %>%
  data.frame(date=time(.), .) %>%
  select_("date", segrun) %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")
tempq_ts <- as.ts(x, start = start(x), end = end(x))

# create monthly model
mmod <- seas(tempm_ts, 
           forecast.maxlead = 30,# extends 30 quarters ahead
           seats.appendfcst="yes",
           # adding this resulted in multiplicative factors
           # I'm not sure why that's the case, without I was 
           # getting additive factors when running on quarterly data
           transform.function = "log") 

# create quarterly model
qmod <- seas(tempq_ts, 
           forecast.maxlead = 30,# extends 30 quarters ahead
           seats.appendfcst="yes",
           # adding this resulted in multiplicative factors
           # I'm not sure why that's the case, without I was 
           # getting additive factors when running on quarterly data
           transform.function = "log") 

# use monthly to create data frame
tempm <- merge(
   # creates and merges invididual zoo series
   oa1 = as.zoo(original(mmod)),
   sa1 = as.zoo(final(mmod)),
   sa1_sf = as.zoo(series(mmod, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.)))

# use quarterly to create data frame
tempq <- merge(
   # creates and merges invididual zoo series
   oa1 = as.zoo(original(qmod)),
   sa1 = as.zoo(final(qmod)),
   sa1_sf = as.zoo(series(qmod, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.)))

```

### Monthly series and seasonal factors: Table  

```{r}
# look at monthly
a <- tempm %>%
  data.frame(date=time(.), .) %>%
  filter(date <= "2015-09-01")
kable(tail(a, n=12), digits = 4, row.names=FALSE)

```

### Monthly series and seasonally adjusted series: Graph  
```{r}
# graph monthly
tempm_graph <- tempm %>%
  data.frame(date=as.Date(time(.)), .) %>%
  gather("variable", "value", -date) %>%
  filter(variable %in% c("oa1", "sa1")) %>%
  filter(date <= "2015-09-01")

p1 <- tempm_graph %>%
   ggplot()
p1 <- p1 + geom_line(data = tempm_graph, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40"))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + scale_y_continuous("axis title")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1
```


### Quarterly series and seasonal factors: Table  
  
```{r}
# look at quarterly
a <- tempq %>%
  data.frame(date=time(.), .) %>%
  filter(date < "2015-09-01")
kable(tail(a, n=12), digits = 4, row.names=FALSE)

```

### Quarterly series and seasonally adjusted series: Graph 
  
```{r}
# graph quarterly
tempq_graph <- tempq %>%
  data.frame(date=as.Date(time(.)), .) %>%
  gather("variable", "value", -date) %>%
  filter(variable %in% c("oa1", "sa1")) %>%
  filter(date < "2015-09-01")

p1 <- tempq_graph %>%
   ggplot()
p1 <- p1 + geom_line(data = tempq_graph, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40"))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + scale_y_continuous("axis title")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1

```


### Use quarterly to estimate monthly

Use the seasonally adjusted quarterly data, and the monthly seasonal factors, to estimate monthly un-adjusted series.  
"oa1q" to indicate non-seasonally adjusted data ("oa1") coming from quarterly data ("q").  
  
- oa1q = sa1q * sa1_sf
  
We observe that oa1q is quite similar to oa1, which is the original monthly data. This implies that if we have a reasonable estimate of quarterly data, then we can use monthly seasonal factors to get to accurate monthly estimates.
```{r}

# use quarterly data to estimate monthly
# creates a sequnce of dats
sq <- seq(as.Date(start(tempq)), as.Date(end(tempq), frac = 1), by = "month") %>%
  as.yearmon(.)
# uses that sequence of dates
zmo <- na.approx(tempq, xout = sq, method = "const")

# drop the original series and seasonal factors
zmo <- zmo %>%
  data.frame(date=as.Date(time(.)), .) %>%
  select(-oa1, -sa1_sf) %>%
  rename(sa1q = sa1) %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() 

tempm1 <- merge(tempm, zmo)

# create a monthly series that is calculated as the
# sa-adjusted quarterly data times the monthly seasonal
# factors
tempm1 <- tempm1 %>%
  transform(oa1q = sa1q * sa1_sf) %>%
  transform(oa1_pchya = oa1 / lag(oa1, 12) - 1) %>%
  transform(oa1q_pchya = oa1q / lag(oa1q, 12) - 1) %>%
  xts()

# look at data
a <- tempm1 %>%
  data.frame(date=time(.), .) %>%
  filter(date < "2015-09-01") %>%
  mutate_each(funs(sprintf("%1.2f%%",100*.)), contains("_pchya"))
kable(tail(a, n=12), digits = 3, row.names=FALSE, align = c("l", rep("r", 7)))
```


### Monthly data, comparison between original and oa1q
```{r}
# graph data
temp_graph <- tempm1 %>%
  data.frame(date=as.Date(time(.)), .) %>%
  gather("variable", "value", -date) %>%
  filter(variable %in% c("oa1", "sa1", "oa1q")) %>%
  filter(date < "2015-09-01" & date > "2013-01-01")

p1 <- temp_graph %>%
   ggplot()
p1 <- p1 + geom_line(data = temp_graph, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40", mypallete[7:7]))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + scale_y_continuous("axis title")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1



```


### Monthly data, comparison between original and oa1q
Percentage change from prior year  
```{r}
# look at it on a percentage change from prior year basis
temp_graph <- tempm1 %>%
  data.frame(date=as.Date(time(.)), .) %>%
  #mutate_each(funs(pchya = . / lag(., 12) - 1), -date) %>%
  gather("variable", "value", -date) %>%
  filter(variable %in% c("oa1_pchya", "oa1q_pchya")) %>%
  filter(date < "2015-09-01" & date > "2010-01-01")

p1 <- temp_graph %>%
   ggplot()
p1 <- p1 + geom_line(data = temp_graph, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40", mypallete[7:7]))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 +  scale_y_continuous(labels=percent)
p1

write.zoo(tempm1, file="output_data/tempm1.csv", sep=",")
```


### Regression, monthly levels  
oa1q explains almost all variation in oa1
```{r }

lmmod_1 <- lm(oa1 ~ oa1q, data = tempm1)
summary(lmmod_1)
# need to set chunk option as results='asis' if I want this table to appear
#tab <- xtable(summary(lmmod_1)$coef, digits=c(0, 2, 2, 1, 2))
#print(tab, type="html")


```


### Regression, monthly percentage change from prior year  
oa1q explains a large proportion of the variation in oa1
```{r}
lmmod_2 <- lm(oa1_pchya ~ oa1q_pchya, data = tempm1)
summary(lmmod_2)
#plot(lmmod_2)
#tab <- xtable(summary(lmmod_2)$coef, digits=c(0, 2, 2, 1, 2))
#print(tab, type="html")

```

