---
title: "Seasonal adjustment comparison"
author: "Tourism Economics"
date: "January 9, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
# require(directlabels)
# require(rmarkdown)
# require(gridExtra)
# require(gtable)
# require(Cairo)

library(arlodr)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
library(seasonal, warn.conflicts=FALSE)
require(xtable)
require(ggplot2)
require(knitr)
require(RColorBrewer)
require(scales)
library(lubridate)

# If I put autodep=TRUE as an argument in the following then I get
# some issues with the graphs, such as grey backgrounds and changes
# in labeling. Maybe it somehow impacts the theme?
#knitr::opts_chunk$set(cache=TRUE, cache.path='output_data/cache/', dpi=96)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, cache.path='../output_data/cache/', autodep=TRUE, fig.path='../output_data/figure_exp_seasonal_short/fig-',  fig.show='asis', fig.keep='high')
# tells it to keep going even if there is an error
knitr::opts_chunk$set(error = TRUE)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("~/Project/R projects/lodfor/output_data/ushist_m.Rdata")
load("~/Project/R projects/lodfor/output_data/ushist_q.Rdata")

```

```{r define_theme, echo=FALSE, message=FALSE, warning=FALSE}
#Displays palette and sets the ts1 theme

#display.brewer.pal(10, "RdBu")
mypallete <- brewer.pal( 10 , "RdBu" )
colors_ts1 <- scale_colour_manual(values = mypallete[c(10, 1, 5, 4)])
# I don't use the following, was an example I saw somewhere maybe as a way to 
# interpolate additional palette colors
#mypal <- colorRampPalette( brewer.pal( 10 , "RdBu" ) )
theme_ts1 <- function(base_size = 12, base_family = "") {
  theme_classic(base_size = base_size) %+replace%
    theme(                                 
      panel.grid.major.x=element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.minor.y=element_blank(),
      axis.ticks.y=element_line(size=.2),
      axis.ticks.x=element_line(size=.2),
      panel.background=element_blank(),
      legend.title=element_blank(),
      legend.key=element_rect(fill="white", colour = "white"),
      legend.key.size=unit(1, "cm"),
      legend.text=element_text(size=rel(1)),
      #legend.position = c(1, .1),
      legend.justification = "right",
      axis.line=element_line(size=.2),
      axis.title.y=element_blank(),
      axis.title.x=element_blank(),
      axis.text=element_text(color="black",size=rel(.7)),
      # following seems to adjust y-axis title
      plot.title=element_text(size=base_size * .8, face="plain",hjust=0,vjust=1)
    #theme(plot.title = element_text(lineheight=.8, face="bold"))
    )
}
theme_set(theme_ts1())
```

```{r}

segrun <- c("totus_demd")

```


### Set up data to run using: `r segrun`
```{r echo=FALSE, message=FALSE, warning=FALSE}

# shorten dates
#ushist_m_1 <- ushist_m["1987-01-01/2015-09-01"]

# sets up a single series to use for a few examples
demo_df <- ushist_m %>%
  data.frame(date=time(.), .) %>%
  select_("date", segrun)

```
  
use default setting
```{r}

# Uses default settings. For example, includes a term for Easter.
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# uses default settings
m1 <- seas(temp_ser_ts, 
           forecast.maxlead = 30,# extends 30 quarters ahead
           seats.appendfcst="yes") 
#summary(m1)

# extract series
temp_sersa1 <- merge(
   # creates and merges invididual zoo series
   oa1 = as.zoo(original(m1)),
   sa1 = as.zoo(final(m1)),
   sa1_sf = as.zoo(series(m1, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.))) %>%
   data.frame(date=as.Date(time(.)), ., row.names=NULL)
```


use default settings m2
```{r}
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# uses default settings
m2 <- seas(temp_ser_ts,
           regression.variables = c("tdnolpyear", "easter[1]", "labor[3]", "ls2001.Sep"),
           forecast.maxlead = 30,# extends 30 quarters ahead
           seats.appendfcst="yes")
# Had been throwing an NaN argument error. I dropped the regression.usertype =
# c("holiday") part because I wasn't using it (wasn't supplying a user
# regressor). Once I dropped it, it works fine.
#summary(m2)

# extract series
temp_sersa2 <- merge(
   # creates and merges invididual zoo series
   oa2 = as.zoo(original(m2)),
   sa2 = as.zoo(final(m2)),
   sa2_sf = as.zoo(series(m2, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.))) %>%
   data.frame(date=as.Date(time(.)), ., row.names=NULL)

```


```{r}
# try setting up several user regression variables as part of a holiday group

x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# days
date <- seq(as.Date("1980-01-01"), as.Date("2030-01-01"), by = 1)
df <- data.frame(date)

# uses lubridate

df1 <- df %>%
  # uses wday from lubridate
  mutate(wday = wday(date, label=TRUE)) %>%
  # grepl returns a logical vector, here it's converted to numeric
  mutate(satsun = as.numeric(grepl("Sat|Sun", wday))) %>%
  # add columns of month day
  mutate(month = months(date)) %>%
  mutate(day = day(date)) %>%
  # identify veterens day
  mutate(vet = ifelse(month == "November" & day == 11, 1, 0)) %>%
  #change values based on a elseif
  mutate(vet_wkend = ifelse(vet == 1 & satsun == 1, 1, 0)) %>%
  #summarise(sum(vet), sum(vet_wkend)) 
  mutate(chr = ifelse(month == "December" & day == 25, 1, 0)) %>%
  #change values based on a elseif
  mutate(chr_wkend = ifelse(chr == 1 & satsun == 1, 1, 0)) %>%
  # July 4
  mutate(jlf = ifelse(month == "July" & day == 4, 1, 0)) %>%
  #change values based on a elseif
  mutate(jlf_wkend = ifelse(jlf == 1 & satsun == 1, 1, 0)) %>%
  # Halloween
  mutate(hlw = ifelse(month == "October" & day == 31, 1, 0)) %>%
  #change values based on a elseif
  mutate(hlw_wkend = ifelse(hlw == 1 & satsun == 1, 1, 0))

# there are a few potential ways to proceed
# create a vector of dates on which veterans day is on a weekend
vet_wkend <- df1 %>%
  filter(vet_wkend == 1)
# use that vector to generate monthly holiday vector for seas
# is a one in November when veterans day is on a weekend
vet_wkend <- vet_wkend$date
vet1 <- genhol(vet_wkend, start = 0, end = 0, frequency = 12)
#vet1 <- window(vet1, start=c(1987, 1), end=c(2030, 12))

chr_wkend <- df1 %>%
  filter(chr_wkend == 1)
chr_wkend <- chr_wkend$date
chr1 <- genhol(chr_wkend, start = 0, end = 0, frequency = 12)
#chr1 <- window(chr1, start=c(1987, 1), end=c(2030, 12)) 

jlf_wkend <- df1 %>%
  filter(jlf_wkend == 1)
jlf_wkend <- jlf_wkend$date
jlf1 <- genhol(jlf_wkend, start = 0, end = 0, frequency = 12)
#jlf1 <- window(jlf1, start=c(1987, 1), end=c(2030, 12)) 

hlw_wkend <- df1 %>%
  filter(hlw_wkend == 1)
hlw_wkend <- hlw_wkend$date
hlw1 <- genhol(hlw_wkend, start = 0, end = 0, frequency = 12)
#hlw1 <- window(hlw1, start=c(1987, 1), end=c(2030, 12)) 

data(holiday)
ea1 <- genhol(easter, start = -2, end = 0, frequency = 12)
#ea1 <- window(ea1, start=c(1987, 1), end=c(2030, 12)) 


# with more regressors
m3 <- seas(x = temp_ser_ts,
           regression.variables = c("tdnolpyear", "labor[3]", "easter[1]", "ls2001.Sep", "thank[3]"),
           forecast.maxlead = 30,# extends 30 quarters ahead
           seats.appendfcst="yes",
           xreg = cbind(vet1, chr1, jlf1), 
           regression.usertype = c("holiday", "holiday2", "holiday3"),
           regression.aictest = NULL)
#summary(m3)


# The basic model used by X-13ARIMA-SEATS for Easter and Labor Day effects assumes that the level of activity changes on the w−th day before the holiday for a specified w, and remains at the new level until the day before the holiday

# plot(m3)
# plot(final(m3))

# extract series
temp_sersa3 <- merge(
   # creates and merges invididual zoo series
   oa3 = as.zoo(original(m3)),
   sa3 = as.zoo(final(m3)),
   sa3_sf = as.zoo(series(m3, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.))) %>%
   data.frame(date=as.Date(time(.)), ., row.names=NULL)

```


```{r}

## Set up Easter
#ea1 is "1" in the month containing the Easter effect, which is set to start two days before Easter

# ea1_df <- ea1 %>% 
#   as.xts(., start = start(.), end = end(.)) %>%
#   data.frame(date=as.Date(time(.)), ., row.names=NULL)
# names(ea1_df) <- c("date", "ea1")
# 
# a <- ea1_df  %>%
#   filter(ea1 == 1) 
# kable(tail(a, n=12), digits = 4, row.names=FALSE)
```

## Set up Veteran's Day weekend
vet_wkend is "1" if Veteran's Day falls on a weekend
```{r}

a <- df1 %>%
  filter(vet == 1) %>%
  select(date:day, vet, vet_wkend)
kable(tail(a, n=12), digits = 4, row.names=FALSE)
```


## Set up Christmas weekend
chr_wkend is "1" if Christmas Day falls on a weekend
```{r}

a <- df1 %>%
  filter(chr == 1) %>%
  select(date:day, chr, chr_wkend)
kable(tail(a, n=12), digits = 4, row.names=FALSE)
```


## Set up July 4th weekend
jlf_wkend is "1" if Independence Day falls on a weekend
```{r}

a <- df1 %>%
  filter(jlf == 1) %>%
  select(date:day, jlf, jlf_wkend)
kable(tail(a, n=12), digits = 4, row.names=FALSE)
```



## Set up three models
m1: defaults  
  
m2: built-in regressors:  
  - "tdnolpyear" (trading days without assuming leap year)   
  - "easter[1]" (Easter, start the effect one day before)  
  - "labor[3]" (Labor Day, start the effect three days before)  
  
m3: same built-in regressors, plus: 
- thank[3]: built-in regressor, number of days in period starting three days before Thanksgiving  
- ls2001.Sep: Assume Sep 11, 2001 is an outlier  
- xreg1: (vet1) Veteran's Day on a weekend  
- xreg2: (chr1) Christmas Day on a weekend  
- xreg3: (jlf1) Independence Day on a weekend  
  

### Original and adjusted series from the models
```{r}

combined <- right_join(temp_sersa1, 
                  temp_sersa2, by = "date")
combined <- right_join(combined, temp_sersa3, by="date")

a <- combined %>%
  filter(date <= "2015-09-01") %>%
  select(date, oa1, oa2, oa3, sa1_sf, sa2_sf, sa3_sf, sa1, sa2, sa3)
kable(tail(a, n=12), digits = 4, row.names=FALSE)

```


### m1
Default
```{r}
summary(m1)
```


### m2
Built-in regressors
```{r}
summary(m2)
```


### m3
Same built-in regressors as m2, plus additional terms
```{r}
summary(m3)
```


### Compare seasonally adjusted series from m1 and m2
Note estimate for Sep. 2015. m1 model tends to indicate Sep. was a "strong" month.  
While m2 model indicates some of that strength was holiday timing. After adjusting, underlying performance was not as strong. Said differently, we would have a high seasonal factor for Sep. 2015. We would divide the unadjusted data by that seasonal factor to get the seasonally adjusted data, which wouldn't be as strong.  
```{r}
temp <- combined %>%
   gather("variable", "value", -date) %>% 
   filter(date >= as.Date("2012-01-01"), date <= as.Date("2015-09-01")) %>%
  filter(variable %in% c("sa1", "sa2"))

p1 <- temp %>%
   ggplot() +
   geom_line(data = temp, aes(x = date, y = value, color=variable), size = .9) +
   scale_color_manual(values = c(mypallete[3:3], "grey40")) +
   labs(colour = "", x = "") +
   scale_y_continuous("axis title") +
   ggtitle("Demand, average daily") +
   theme(legend.position = "bottom")
p1

```


### Compare seasonally adjusted series from m1 and m3
Again, similar to m2, m3 indicates Sep. 2015 was not as strong as m1 results would indicate.
```{r}
temp <- combined %>%
   gather("variable", "value", -date) %>% 
   filter(date >= as.Date("2012-01-01"), date <= as.Date("2015-09-01")) %>%
  filter(variable %in% c("sa1", "sa3"))

p1 <- temp %>%
   ggplot() +
   geom_line(data = temp, aes(x = date, y = value, color=variable), size = .9) +
   scale_color_manual(values = c(mypallete[3:3], "grey40")) +
   labs(colour = "", x = "") +
   scale_y_continuous("axis title") +
   ggtitle("Demand, average daily") +
   theme(legend.position = "bottom")
p1

```


### Compare seasonally adjusted series from m2 and m3
Both m2 and m3 are similar for Sep. 2015. Interestingly, m3 generally points to performance during the summer as being somewhat weaker than m2.  

```{r}
temp <- combined %>%
   gather("variable", "value", -date) %>% 
   filter(date >= as.Date("2012-01-01"), date <= as.Date("2015-09-01")) %>%
  filter(variable %in% c("sa2", "sa3"))

p1 <- temp %>%
   ggplot() +
   geom_line(data = temp, aes(x = date, y = value, color=variable), size = .9) +
   scale_color_manual(values = c(mypallete[3:3], "grey40")) +
   labs(colour = "", x = "") +
   scale_y_continuous("axis title") +
   ggtitle("Demand, average daily") +
   theme(legend.position = "bottom")
p1

```


### Seasonal factors for a particular month: Sep. Labor Day timing

Year   |   Labor Day date    
-------|----------
2011 | 5  
2012 | 3  
2013 |  2  
2014 |  1  
2015 | 7  
2016 |5  


### Seasonal factors for September  
  
```{r}

a <- combined %>%
  filter(date <= "2016-09-01") %>%
  mutate(month = lubridate::month(date, label=TRUE)) %>%
  filter(month == "Sep") %>%
  select(date, sa1_sf, sa2_sf, sa3_sf)
kable(tail(a, n=12), digits = 4, row.names=FALSE)

```


### Seasonal factors for September (1987 to 2016)
Last point is Sep 2016. So both Sep. 2015 and Sep. 2016 have large seasonal factors. 
```{r}
temp <- combined %>%
  filter(date <= as.Date("2016-09-01")) %>%
  mutate(month = lubridate::month(date, label=TRUE)) %>%
  filter(month == "Sep") %>%
  select(date, sa3_sf) %>%
  gather("variable", "value", -date) 

p1 <- temp %>%
   ggplot() +
   geom_line(data = temp, aes(x = date, y = value, color=variable), size = .9) +
   scale_color_manual(values = c(mypallete[3:3], "grey40")) +
   labs(colour = "", x = "") +
   scale_y_continuous("axis title") +
   ggtitle("Seasonal factor") +
   theme(legend.position = "bottom")
p1 <- p1 + scale_x_date(date_breaks = "1 year", date_labels = "%y")
p1

```

