---
title: "Create seasonal dummy and dates"
author: "Tourism Economics"
date: "January 9, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(arlodr)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(readr, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
library(seasonal, warn.conflicts=FALSE)
require(xtable)
require(ggplot2)
require(knitr)
require(RColorBrewer)
require(scales)
library(lubridate)

# If I put autodep=TRUE as an argument in the following then I get
# some issues with the graphs, such as grey backgrounds and changes
# in labeling. Maybe it somehow impacts the theme?
#knitr::opts_chunk$set(cache=TRUE, cache.path='output_data/cache/', dpi=96)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, cache.path='../output_data/cache/', autodep=TRUE, fig.path='../output_data/figure_exp_seasonal_short/fig-',  fig.show='asis', fig.keep='high')
# tells it to keep going even if there is an error
knitr::opts_chunk$set(error = TRUE)

```


The goal of this script is to create a dataframe with a useful set of holiday regressors. 

I originally approached it by creating a date vector for each holiday. But moving those around in lists got difficult. And I switched to creating a single daily data frame as the export

As a first step, I'll handle some easy holidays that are on the same date each year.

```{r}

# days
date <- seq(as.Date("1970-01-01"), as.Date("2030-01-01"), by = 1)
df <- data.frame(date)

same_date <- df %>%
  # add columns of month day
  mutate(month = months(date)) %>%
  mutate(day = day(date)) %>%
  # Valentine's day
  mutate(val = ifelse(month == "February" & day == 14, 1, 0)) %>%
  # July 4
  mutate(jlf = ifelse(month == "July" & day == 4, 1, 0)) %>%
  # veterens day
  mutate(vet = ifelse(month == "November" & day == 11, 1, 0)) %>%
   # Halloween
  mutate(hlw = ifelse(month == "October" & day == 31, 1, 0)) %>%
  # christmas
  mutate(chr = ifelse(month == "December" & day == 25, 1, 0)) %>%
  # new years
  mutate(nyd = ifelse(month == "January" & day == 1, 1, 0))

# 
# # creates a vector of each
# val <- same_date %>%
#   filter(val == 1) %>%
#   # creates vector of the date column
#   .[,'date']
# 
# jlf <- same_date %>%
#   filter(jlf == 1) %>%
#   # creates vector of the date column
#   .[,'date']
# 
# vet <- same_date %>%
#   filter(vet == 1) %>%
#   # creates vector of the date column
#   .[,'date']
# 
# hlw <- same_date %>%
#   filter(hlw == 1) %>%
#   # creates vector of the date column
#   .[,'date']
# 
# chr <- same_date %>%
#   filter(chr == 1) %>%
#   # creates vector of the date column
#   .[,'date']
# 
# nyd <- same_date %>%
#   filter(chr == 1) %>%
#   # creates vector of the date column
#   .[,'date']

```


Import some csv files that I manually created based on the timeanddate website. Pretty easy.

```{r}

# days
date <- seq(as.Date("1970-01-01"), as.Date("2030-01-01"), by = 1)
df <- data.frame(date)

fpath <- c("~/Project/R projects/lodfor")

hold_df <- df

# president's day
fname <- paste0(fpath, "/input_data/timeanddate_pres_us.csv")
in_1 <- read_csv(fname, col_names=TRUE) 
colnames(in_1) <- colnames(in_1) %>%
  tolower() %>%
  gsub("[[:punct:]]", "_", .) %>%
  gsub(" ", "_", .)
in_2 <- in_1 %>%
  mutate(date = paste(date, year, sep="-")) %>%
  mutate(date = as.Date(date, "%d-%b-%Y")) %>%
  select(date) %>%
  # name the holiday
  mutate(prs = 1) %>%
  left_join(df, ., by="date") %>%
  replace(is.na(.), 0) 
hold_df <- merge(hold_df, in_2, by="date")

# memorial day, started as Federal holiday on final monday in 1971
# so I trimmed it
fname <- paste0(fpath, "/input_data/timeanddate_mem_us.csv")
in_1 <- read_csv(fname, col_names=TRUE) 
colnames(in_1) <- colnames(in_1) %>%
  tolower() %>%
  gsub("[[:punct:]]", "_", .) %>%
  gsub(" ", "_", .)
# added to handle memorial day
in_1 <- in_1 %>%
  filter(holiday_type == "Federal Holiday")
in_2 <- in_1 %>%
  mutate(date = paste(date, year, sep="-")) %>%
  mutate(date = as.Date(date, "%d-%b-%Y")) %>%
  select(date) %>%
  # name the holiday
  mutate(mem = 1) %>%
  left_join(df, ., by="date") %>%
  replace(is.na(.), 0) 
hold_df <- merge(hold_df, in_2, by="date")


# easter
fname <- paste0(fpath, "/input_data/timeanddate_easter_us.csv")
in_1 <- read_csv(fname, col_names=TRUE) 
colnames(in_1) <- colnames(in_1) %>%
  tolower() %>%
  gsub("[[:punct:]]", "_", .) %>%
  gsub(" ", "_", .)
in_2 <- in_1 %>%
  mutate(date = paste(date, year, sep="-")) %>%
  mutate(date = as.Date(date, "%d-%b-%Y")) %>%
  select(date) %>%
  # name the holiday
  mutate(eas = 1) %>%
  left_join(df, ., by="date") %>%
  replace(is.na(.), 0) 
hold_df <- merge(hold_df, in_2, by="date")

# Presidential election days
fname <- paste0(fpath, "/input_data/timeanddate_electionday_us.csv")
in_1 <- read_csv(fname, col_names=TRUE) 
colnames(in_1) <- colnames(in_1) %>%
  tolower() %>%
  gsub("[[:punct:]]", "_", .) %>%
  gsub(" ", "_", .)
in_2 <- in_1 %>%
  mutate(date = paste(date, year, sep="-")) %>%
  mutate(date = as.Date(date, "%d-%b-%Y")) %>%
  select(date) %>%
  # name the holiday
  mutate(ele = 1) %>%
  left_join(df, ., by="date") %>%
  replace(is.na(.), 0) 
hold_df <- merge(hold_df, in_2, by="date")

# Thanksgiving
fname <- paste0(fpath, "/input_data/timeanddate_thanksgiving_us.csv")
in_1 <- read_csv(fname, col_names=TRUE) 
colnames(in_1) <- colnames(in_1) %>%
  tolower() %>%
  gsub("[[:punct:]]", "_", .) %>%
  gsub(" ", "_", .)
in_2 <- in_1 %>%
  mutate(date = paste(date, year, sep="-")) %>%
  mutate(date = as.Date(date, "%d-%b-%Y")) %>%
  select(date) %>%
  # name the holiday
  mutate(thk = 1) %>%
  left_join(df, ., by="date") %>%
  replace(is.na(.), 0) 
hold_df <- merge(hold_df, in_2, by="date")

# Hanukkah
fname <- paste0(fpath, "/input_data/timeanddate_hanukkahfirstd_us.csv")
in_1 <- read_csv(fname, col_names=TRUE) 
colnames(in_1) <- colnames(in_1) %>%
  tolower() %>%
  gsub("[[:punct:]]", "_", .) %>%
  gsub(" ", "_", .)
in_2 <- in_1 %>%
  mutate(date = paste(date, year, sep="-")) %>%
  mutate(date = as.Date(date, "%d-%b-%Y")) %>%
  select(date) %>%
  # name the holiday
  mutate(han = 1) %>%
  left_join(df, ., by="date") %>%
  replace(is.na(.), 0) 
hold_df <- merge(hold_df, in_2, by="date")



```


```{r}

# so what if I then want to put these into a dataframe together

hol_df1 <- merge(same_date, hold_df, by="date")

# # days
# date <- seq(as.Date("1980-01-01"), as.Date("2030-01-01"), by = 1)
# df <- data.frame(date)
# 
# # create a dataframe of each one
# val_df <- val %>%
#   data.frame(date = .) %>%
#   mutate(val = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# eas_df <- eas %>%
#   data.frame(date = .) %>%
#   mutate(eas = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# mem_df <- mem %>%
#   data.frame(date = .) %>%
#   mutate(mem = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# jlf_df <- jlf %>%
#   data.frame(date = .) %>%
#   mutate(jlf = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# hlw_df <- hlw %>%
#   data.frame(date = .) %>%
#   mutate(hlw = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# ele_df <- ele %>%
#   data.frame(date = .) %>%
#   mutate(ele = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# vet_df <- vet %>%
#   data.frame(date = .) %>%
#   mutate(vet = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# thk_df <- thk %>%
#   data.frame(date = .) %>%
#   mutate(thk = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# han_df <- han %>%
#   data.frame(date = .) %>%
#   mutate(han = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# chr_df <- chr %>%
#   data.frame(date = .) %>%
#   mutate(chr = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)
# 
# nyd_df <- nyd %>%
#   data.frame(date = .) %>%
#   mutate(nyd = 1) %>%
#   left_join(df, ., by="date") %>%
#   replace(is.na(.), 0)

# # combine by creating a list and merging that list
# hol_df1 <- Reduce(function(x, y) merge(x, y, all=TRUE), 
#                   list(val_df, eas_df, mem_df, jlf_df, hlw_df,
#                        ele_df, vet_df, thk_df, han_df, chr_df,
#                        nyd_df))

# create a dataframe with various weekday and weekend flags
date <- seq(as.Date("1970-01-01"), as.Date("2030-01-01"), by = 1)
df <- data.frame(date)

flag_df <- df %>%
  # add columns of month day
  mutate(month = months(date)) %>%
  mutate(day = day(date)) %>%
  # uses wday from lubridate
  mutate(wday = wday(date, label=TRUE)) %>%
  # grepl returns a logical vector, here it's converted to numeric
  mutate(satsun = as.numeric(grepl("Sat|Sun", wday))) %>%
  mutate(frisat = as.numeric(grepl("Fri|Sat", wday))) %>%
  mutate(fss = as.numeric(grepl("Fri|Sat|Sun", wday))) %>%
  mutate(fssm = as.numeric(grepl("Fri|Sat|Sun|Mon", wday)))

# combine
hol_df2 <- merge(flag_df, hol_df1)

```


## What to save
Set up to save a daily data frame with the holidays as dummy variables. Also save the date vectors.
```{r}

holiday_daily <- hol_df2
# saves Rdata version of the data
save(holiday_daily, file="output_data/holiday_daily.Rdata")

# holiday_dates <- list("val"=val, "eas"=eas, "mem"=mem, "jlf"=jlf, "hlw"=hlw,
#                       "ele"=ele, "vet"=vet, "thk"=thk, "han"=han, "chr"=chr, "nyd"=nyd)
# # saves Rdata version of the data
# save(holiday_dates, file="output_data/holiday_dates.Rdata")

```

