---
title: "Create seasonal regressors"
author: "Tourism Economics"
date: "February 12, 2016"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(arlodr)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(readr, warn.conflicts=FALSE)
library(readxl, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
library(seasonal, warn.conflicts=FALSE)
require(xtable)
require(ggplot2)
require(knitr)
require(RColorBrewer)
require(scales)
library(lubridate)
library(lazyeval, warn.conflicts=FALSE)

# If I put autodep=TRUE as an argument in the following then I get
# some issues with the graphs, such as grey backgrounds and changes
# in labeling. Maybe it somehow impacts the theme?
#knitr::opts_chunk$set(cache=TRUE, cache.path='output_data/cache/', dpi=96)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, cache.path='../output_data/cache/', autodep=TRUE, fig.path='../output_data/figure_exp_seasonal_short/fig-',  fig.show='asis', fig.keep='high')
# tells it to keep going even if there is an error
knitr::opts_chunk$set(error = TRUE)

```

Load the files containing the dates of the holidays. 
Loads a daily data frame with the holidays as dummy variables. 

```{r}
load(file="output_data/holiday_daily.Rdata")

# originally I had a list of vectors of holiday dates that I was opening. I
# used this to turn a list in which each is a character vector, and turned it into 
# a dataframe
# dl <- data.frame(h_code = rep(names(holiday_dates), sapply(holiday_dates, length)),
#                  date = unlist(holiday_dates)) %>%
#   mutate(date = as.Date(date))

# load in daily data used to estimate typical day of week averages for certain months. 
# This file doesn't have to be kept up to date, since we're using longer run averages.
usdaily <- read_excel(path="input_data/TEDailyUS_201509.xls", col_names=TRUE)
colnames(usdaily) <- tolower(colnames(usdaily))
usdaily <- usdaily %>%
  mutate(date = as.Date(as.character(date), "%Y%m%d"))
```

```{r}
# calculate certain day of week averages on a long-term basis

d_avg <- usdaily %>%
  mutate(month = months(date)) %>% 
  mutate(day = day(date)) %>%
  mutate(wday = wday(date, label=TRUE)) %>%
  mutate(demand_roll = rollapply(data = demand, 
                            width = 28, 
                            FUN = mean, 
                            align = "right", 
                            fill = NA, 
                            na.rm = T)) %>%
  mutate(ratio_roll = demand/demand_roll)

```



```{r}

# This is a bulk approach that creates date vectors for each of the holidays in
# the holiday_daily data frame. These can then be used in the individual holiday
# work below. I'll need to see if I use this approach, or if it might be fine to
# just do it one by one when working through each holiday.

# gather the data frame so that the holiday codes are in a single column
# and the dates are in a column
hold_vec <- holiday_daily %>%
  select(date, val:han) %>%
  gather(h_code, value, -date) %>%
  filter(value == 1)

# this link was helpful as I wrote the following
#https://codedocean.wordpress.com/2014/01/14/r-working-with-named-objects-in-a-loop/
varNames <- as.character(unique(hold_vec$h_code))
for (i in seq_along(varNames)) {
 hold_a <- filter(hold_vec, h_code == varNames[i]) %>%
 # creates a vector
  .[,'date']
 # names the vector, this uses the first object as a name for the second object
 assign(varNames[i], hold_a)
}

# as an example, then I can use the resulting vectors of dates in genhol
temp_ts_m <- genhol(thk, start = 0, end = 6, frequency=12)
temp_ts_m

```



```{r}

# some of the regressors are created by constructing certain situations

df1 <- holiday_daily %>%
  mutate(vet_wkend = ifelse(vet == 1 & satsun == 1, 1, 0)) %>%
  mutate(chr_wkend = ifelse(chr == 1 & satsun == 1, 1, 0)) %>%
  mutate(jlf_wkend = ifelse(jlf == 1 & satsun == 1, 1, 0)) %>%
  mutate(hlw_wkend = ifelse(hlw == 1 & satsun == 1, 1, 0))

vet_wkend <- df1 %>%
  filter(vet_wkend == 1)

# use that vector to generate monthly holiday vector for seas
# is a one in November when veterans day is on a weekend
vet_wkend <- vet_wkend$date
vet1_ts_m <- genhol(vet_wkend, start = 0, end = 0, frequency = 12)
vet1_ts_m

chr_wkend <- df1 %>%
  filter(chr_wkend == 1)
chr_wkend <- chr_wkend$date
chr1 <- genhol(chr_wkend, start = 0, end = 0, frequency = 12)
#chr1 <- window(chr1, start=c(1987, 1), end=c(2030, 12)) 

jlf_wkend <- df1 %>%
  filter(jlf_wkend == 1)
jlf_wkend <- jlf_wkend$date
jlf1 <- genhol(jlf_wkend, start = 0, end = 0, frequency = 12)
#jlf1 <- window(jlf1, start=c(1987, 1), end=c(2030, 12)) 

hlw_wkend <- df1 %>%
  filter(hlw_wkend == 1)
hlw_wkend <- hlw_wkend$date
hlw1 <- genhol(hlw_wkend, start = 0, end = 0, frequency = 12)
#hlw1 <- window(hlw1, start=c(1987, 1), end=c(2030, 12)) 


```

```{r}

# Presumably there could be another approach that doesn't use genhol. The end goal is to get to a time series object that can be used as a regressor.  

```


## Holiday backgrounds and thoughts

### Valentine's Day
When: February 14
Discussion comments: When it falls on a Saturday it is huge. Falling on Friday was good, but wasn’t great. How does it intereact with Presidents' day (e.g. if it is on the Sunday before President’s day)?

```{r}

df1 <- holiday_daily %>%
  select(date:fssm, val, prs) %>%
  mutate(val_fss = ifelse(val == 1 & fss == 1, 1, 0)) %>%
  mutate(val_fs = ifelse(val == 1 & frisat == 1, 1, 0)) %>%
  mutate(val_fri = ifelse(val == 1 & wday == "Fri", 1, 0)) %>%
  mutate(val_sat = ifelse(val == 1 & wday == "Sat", 1, 0)) %>%
  # creates dummy that is one the day before presidents day
  mutate(prs_lead1 = lead(prs, order_by=date)) %>%
  # two days before
  mutate(prs_lead2 = lead(prs, n = 2, order_by=date)) %>%
  # val_prs will be one if one of the two other dummies is one
  # so it is a one if valentine's day is a saturday or sunday of president's day weekend
  mutate(val_prs = (prs_lead1 + prs_lead2) * val)

# take a look
a <- df1 %>%
  filter(val == 1) %>%
  select(date:day, wday, val, val_fss, val_fs, val_fri, val_sat, val_prs)
kable(tail(a, n=12), digits = 4, row.names=FALSE)

# filter and use that vector to generate monthly holiday vector 
# val_fss <- df1 %>%
#   filter(val_fss == 1) 
# val_fss <- val_fss$date

# bulk approach of creating vectors based on initial example
hold_vec <- df1 %>%
  # set this part specifically, rest should be fine
  select(date, val_fss:val_prs)  %>%
  gather(h_code, value, -date) %>%
  filter(value == 1)
varNames <- as.character(unique(hold_vec$h_code))
for (i in seq_along(varNames)) {
 hold_a <- filter(hold_vec, h_code == varNames[i]) %>%
 # creates a vector
  .[,'date']
 # names the vector, this uses the first object as a name for the second object
 assign(varNames[i], hold_a)
}

# use genhol to generate time series object
val_fss_ts_m <- genhol(val_fss, start = 0, end = 0, frequency = 12)
val_fs_ts_m <-  genhol(val_fs, start = 0, end = 0, frequency = 12)
val_fri_ts_m <- genhol(val_fri, start = 0, end = 0, frequency = 12)
val_sat_ts_m <- genhol(val_sat, start = 0, end = 0, frequency = 12)
val_prs_ts_m <- genhol(val_prs, start = 0, end = 0, frequency = 12)

v_names <- c("val_fss_ts_m", "val_fs_ts_m", "val_fri_ts_m", "val_sat_ts_m", "val_prs_ts_m")

val_list <- create_named_list(v_names)
#val_fri_ts_m
#val_fs_ts_m
```



### Easter
When: Many Christians celebrate Jesus Christ's resurrection on Easter Sunday. The Easter date depends on the ecclesiastical approximation of the March equinox. Palm Sunday is the Sunday before Easter Sunday. Easter can happen as early as March 22, but that is very rare. Looking at the 1950 to 2050 period, Easter happened once as early as March 23, but generally happens between March 25 and April 22. In this period, Easter happened as late as April 25.
Discussion comments: Easter has a prolonged impact. Starts the Monday before, goes to the Monday after. 
Approach: The effect could start as early as the Sunday before Easter. That Sunday might not experience much of an effect. The Easter effect might also extend a day after easter. So I would tend to look at the.


```{r}
df1 <- holiday_daily %>%
  select(date:fssm, eas) 

# take a look
a <- df1 %>%
  filter(eas == 1) %>%
  select(date:day, wday, eas)
kable(tail(a, n=12), digits = 4, row.names=FALSE)

# bulk approach of creating vectors based on initial example
hold_vec <- df1 %>%
  # set this part specifically, rest should be fine
  select(date, eas)  %>%
  gather(h_code, value, -date) %>%
  filter(value == 1)
varNames <- as.character(unique(hold_vec$h_code))
for (i in seq_along(varNames)) {
 hold_a <- filter(hold_vec, h_code == varNames[i]) %>%
 # creates a vector
  .[,'date']
 # names the vector, this uses the first object as a name for the second object
 assign(varNames[i], hold_a)
}

# use genhol to generate time series object
eas_7_1_ts_m <- genhol(eas, start = -7, end = 1, frequency = 12)

v_names <- c("eas_7_1_ts_m")
eas_list <- create_named_list(v_names)

```



### Memorial Day 
When: Last Monday of May
Discussion comments: Sometimes it is the very last bit of May. If you have a full week after Memorial Day that is in May, then it’s a slow week. So if Memorial Day is the last day of may, then you’ve pushed that slowness into June. 

Note: 
Memorial Day was fixed to the last Monday in May for the first time in 1971 based on the input file. So I filtered to only those dates, I think that means 1970 doesn't have a Memorial Day flagged in the holiday_dates table.

Approach for end of May: 
Created regressor to measure how many business days follow Memorial Day, in May. 

Also,created a weighted version of that regressor. In calculating the weighted version, this looks at the typical decline for that specific day of the week. For example, considering just the period after May 15, demand on the typcal Tuesday before Memorial is at 1.08 of the rolling average, while after Memorial Day the 14 sample Tuesdays are at 0.85, so a 21% decline. This is based on the following table.

The approach is to calculate weight that is the average of those percentage declines. That helps gives a sense of whether the days that are following Memorial Day in May tend to be days with higher impacts. 

```{r Memorial Day end of May}

# this dataframe is used for two separate approaches
df1 <- holiday_daily %>%
  select(date:fssm, mem) %>%
  # some temporoary colummns
  mutate(mempost1 = ifelse(lag(mem, n=1, order_by=date) == 1, 1, 0)) %>%
  mutate(mempost2 = ifelse(lag(mem, n=2, order_by=date) == 1, 1, 0)) %>%
  mutate(mempost3 = ifelse(lag(mem, n=3, order_by=date) == 1, 1, 0)) %>%
  mutate(mempost4 = ifelse(lag(mem, n=4, order_by=date) == 1, 1, 0)) %>%
  mutate(mempost5 = ifelse(lag(mem, n=5, order_by=date) == 1, 1, 0)) %>%
  mutate(mempost6 = ifelse(lag(mem, n=6, order_by=date) == 1, 1, 0)) %>%
  # used to calculate an indicator of whether the day is after Memorial day
  mutate(mempost = mempost1 + mempost2 + mempost3 + mempost4 + mempost5 + mempost6) %>%
  # fix days in any month other than May (namely June) to be zero
  mutate(mempost = ifelse(month == "May", mempost, 0)) %>%
  # remove temp columns
  select(-mempost1, -mempost2, -mempost3, -mempost4, -mempost5, -mempost6)

#####
#
# Analysis of Tuesdays-Wednesdays and weekend days that follow Memorial Day
#

# calculate days
mem_buswke <- df1 %>%
  mutate(mempost_tw = ifelse(mempost == 1 & wday == "Tues" , 1, 
                             ifelse(mempost == 1 & wday == "Wed" , 1, 0))) %>%
  mutate(mempost_wke = ifelse(mempost == 1 & fss == 1, 1, 0)) %>%
  mutate(year = year(date)) %>%
  mutate(month = month(date, label = FALSE)) %>%
  select(year, month, mempost_tw, mempost_wke) %>%
  group_by(year, month) %>%
  summarize_each(funs(sum), mempost_tw, mempost_wke) %>%
  ungroup() %>%
  mutate(date = paste(year, month, sep="-")) %>%
  mutate(date = as.yearmon(date)) %>%
  mutate(date = as.Date(date)) %>%
  select(date, mempost_tw, mempost_wke) %>%
  data.frame()

# use function to add a centered regressor
mem_buswke2 <- mem_buswke %>%
  # first argument is the dataframe, second is the column to create the
  # centered regressor of
  arlodr::center_reg("mempost_tw") %>%
  arlodr::center_reg("mempost_wke")

# use a function to convert select series to monthly ts objects
mempost_tw_m <- arlodr::create_ts_reg_m(mem_buswke2, "mempost_tw_cent") 
mempost_wke_m <- arlodr::create_ts_reg_m(mem_buswke2, "mempost_wke_cent")


########
#
# creates regressor based on the average effect of each day

mem1 <- df1 %>%
  left_join(select(d_avg, date, ratio_roll), by = "date") %>%
  filter(ratio_roll > 0 & month == "May" & day > 15 & wday != "Mon") %>%
  group_by(month, wday, mempost) %>%
  summarize(avg = mean(ratio_roll), count = n()) %>%
  mutate(avg_diff  = avg/lag(avg, n=1) - 1) %>%
  select(month:avg, avg_diff, count)
kable(mem1, digits = 2, row.names=FALSE)

# average difference by day of week
mem2 <- mem1 %>%
  filter(mempost == 1) %>%
  select(month, wday, mempost, avg_diff)
kable(mem2, digits = 2, row.names=FALSE)

# appends the daily data by adding the average difference by day of week
df2 <- df1 %>%
  # select the rows from df1 to work with
  filter(month == "May" & mempost == 1) %>%
  # this adds the avg_diff column to the corresponding rows
  left_join(mem2, by = "wday") %>%
  select(date, avg_diff)

df3 <- df1 %>%
  # then add that avg_diff column onto the original df1
  full_join(df2, by = "date") %>%
  # replace na's with zeros
  mutate(avg_diff = ifelse(is.na(avg_diff),0,avg_diff)) %>%
  # calculate weighted mempost as the sum of the avg_dff column for each May
  mutate(year_mon = paste(year(date), month(date), sep="-")) %>%
  group_by(year_mon) %>%
  summarize(mem_diff = sum(avg_diff)) 

# calculate centered regressor
df4 <- df3 %>%
  mutate(date = as.Date(as.yearmon(year_mon))) %>%
  select(-year_mon) %>%
  arlodr::center_reg("mem_diff") %>%
  # keep date column in the front
  select(date, everything()) %>%
  arrange(date) %>%
  data.frame()

# use a function to convert select series to monthly ts objects

mempost_m <- arlodr::create_ts_reg_m(df4, "mem_diff_cent") 

v_names <- c("mempost_tw_m", "mempost_wke_m", "mempost_m")

mem_list <- create_named_list(v_names)

```


### Fourth of July
When: July 4
Discussion comments: Can have a bit of an impact on June. If it is on a Friday for example, then it’s a slower week that last week of June. Whereas, if it’s on Monday or Tuesday, then the June week is still strong. And it’s the week in July that will be week. 

### Labor Day - Start of September
When: First Monday of September
Discussion comments: Septembers that start on a Saturday, vs. Septembers that start on a Wednesday. Could potentially have an impact on August. So if August ends on Saturday, that’s going to make a strong last day of August. Also when August ends on Sunday, it’s pretty strong. 

Approach for end of August:
Created regressors that reflect whether the end of August is on a weekend. 

```{r Labor Day end of August}

df1 <- holiday_daily %>%
  select(date:fssm) %>%
  mutate(augend = ifelse(month == "August" & day == 31, 1, 0)) %>%
  mutate(augend_fss = ifelse(augend == 1 & fss == 1, 1, 0)) %>%
  mutate(augend_ss = ifelse(augend == 1 & satsun == 1, 1, 0)) 

# take a look
a <- df1 %>%
  filter(augend == 1) %>%
  select(date:day, wday, augend, augend_fss, augend_ss)
kable(tail(a, n=12), digits = 4, row.names=FALSE)

# bulk approach of creating vectors based on initial example
hold_vec <- df1 %>%
  # set this part specifically, rest should be fine
  select(date, augend_fss:augend_ss)  %>%
  gather(h_code, value, -date) %>%
  filter(value == 1)
varNames <- as.character(unique(hold_vec$h_code))
for (i in seq_along(varNames)) {
 hold_a <- filter(hold_vec, h_code == varNames[i]) %>%
 # creates a vector
  .[,'date']
 # names the vector, this uses the first object as a name for the second object
 assign(varNames[i], hold_a)
}

# use genhol to generate time series object
augend_fss_ts_m <- genhol(augend_fss, start = 0, end = 0, frequency = 12, center = "calendar")
augend_ss_ts_m <-  genhol(augend_ss, start = 0, end = 0, frequency = 12, center = "calendar")

v_names <- c("augend_fss_ts_m", "augend_ss_ts_m")

augend_list <- create_named_list(v_names)

#####
#
# take a peek at the centered series
#

# create xts objects
augend_fss_x_m <- augend_fss_ts_m %>%
  xts(order.by=as.yearmon(index(.)))
augend_ss_x_m <- augend_ss_ts_m %>%
  xts(order.by=as.yearmon(index(.)))

# create simple table
a <- merge(augend_fss_x_m, augend_ss_x_m) %>%
    data.frame(date=time(.), .) %>%
  mutate(month = month(date)) %>%
  filter(month %in% c(8))
kable(head(a, n=8), digits = c(4, 4, 4), row.names=FALSE)

```

Approach for September
Created regressors that reflect how September starts

```{r Labor Day beg of September}

df1 <- holiday_daily %>%
  select(date:fssm) %>%
  mutate(sepstr = ifelse(month == "September" & day == 1, 1, 0)) %>%
  mutate(sepstr_fssm = ifelse(sepstr == 1 & fssm == 1, 1, 0)) %>%
  mutate(sepstr_ss = ifelse(sepstr == 1 & satsun == 1, 1, 0)) %>%
  mutate(sepstr_twt = ifelse(sepstr == 1 & wday %in% c("Tues", "Wed", "Thurs"), 1, 0)) %>%
  mutate(sepstr_mon = ifelse(sepstr == 1 & wday == "Mon", 1, 0)) %>%
  mutate(sepstr_tue = ifelse(sepstr == 1 & wday == "Tues", 1, 0)) %>%
  mutate(sepstr_wed = ifelse(sepstr == 1 & wday == "Wed", 1, 0)) %>%
  mutate(sepstr_thr = ifelse(sepstr == 1 & wday == "Thurs", 1, 0)) %>%
  mutate(sepstr_fri = ifelse(sepstr == 1 & wday == "Fri", 1, 0))

# take a look
a <- df1 %>%
  filter(sepstr == 1) %>%
  select(date:day, wday, sepstr, sepstr_fssm, sepstr_ss, sepstr_twt, sepstr_mon, sepstr_tue, sepstr_wed, sepstr_thr, sepstr_fri)
kable(head(a, n=12), digits = 4, row.names=FALSE)

# bulk approach of creating vectors based on initial example
hold_vec <- df1 %>%
  # set this part specifically, rest should be fine
  select(date, sepstr_fssm:sepstr_fri)  %>%
  gather(h_code, value, -date) %>%
  filter(value == 1)
varNames <- as.character(unique(hold_vec$h_code))
for (i in seq_along(varNames)) {
 hold_a <- filter(hold_vec, h_code == varNames[i]) %>%
 # creates a vector
  .[,'date']
 # names the vector, this uses the first object as a name for the second object
 assign(varNames[i], hold_a)
}

# use genhol to generate time series object
sepstr_fssm_ts_m <- genhol(sepstr_fssm, start = 0, end = 0, frequency = 12, center = "calendar")
sepstr_ss_ts_m <-  genhol(sepstr_ss, start = 0, end = 0, frequency = 12, center = "calendar")
sepstr_twt_ts_m <-  genhol(sepstr_twt, start = 0, end = 0, frequency = 12, center = "calendar")
sepstr_mon_ts_m <-  genhol(sepstr_mon, start = 0, end = 0, frequency = 12, center = "calendar")
sepstr_tue_ts_m <-  genhol(sepstr_tue, start = 0, end = 0, frequency = 12, center = "calendar")
sepstr_wed_ts_m <-  genhol(sepstr_wed, start = 0, end = 0, frequency = 12, center = "calendar")
sepstr_thr_ts_m <-  genhol(sepstr_thr, start = 0, end = 0, frequency = 12, center = "calendar")
sepstr_fri_ts_m <-  genhol(sepstr_fri, start = 0, end = 0, frequency = 12, center = "calendar")

v_names <- c("sepstr_fssm_ts_m", "sepstr_ss_ts_m", "sepstr_twt_ts_m", "sepstr_mon_ts_m", "sepstr_tue_ts_m", "sepstr_wed_ts_m", "sepstr_thr_ts_m", "sepstr_fri_ts_m")

sepstr_list <- create_named_list(v_names)

```


### Halloween
When: October 31
Discussion comments: Maybe it’s three groups: Sunday-midweek, Friday, Saturday. 

### Presidential election days
When: Election Day in the United States of America is the Tuesday following the first Monday in November. It can fall on or between November 2 and November 8. 

### Veteran’s day
When: November 11
Discussion comments: Minimal impact

### Thanksgiving
When: Fourth Thursday of November
Discussion comments: Does Thanksgiving week end the month, or is there another week after it. How many days are there between Thanksgiving and christmats. If you have an extra week before thanksgiving and December, then you might get a good week that first week of December because you have more time. 

### Hanukkah
When: Hanukkah is an eight-day Jewish observance.The first night of Hanukkah is the night before the first day of Hanukkah. From 1950 to 2050, the earliest is Nov. 28 and the latest is Dec. 27. About a tenth of the time it starts in those last few days of November. About a third of the time it starts on or before Dec. 7. About a fifth of the time it starts on or after Dec. 22. 
Discussion comments: Hanukah doesn’t have a huge impact. Mainly groups and there isn’t that much business in December. 
Sometimes Hanuakah and Christmas intersect

### Christmas
Discussion comments: Christmas on the weekend vs. the middle of the week. 

### New Year's Day

### Leap year
Discussion comments: STR reallocates the day, so take total, divide by 29, then multiply by 28. 


# Save results
I can then save multiple objects into a single file
```{r}
# save, with list of the objects to save together in the Rdata file
save(val_list, mem_list, eas_list, augend_list, sepstr_list, file="output_data/holiday_regress.Rdata")

```
