---
title: "Seasonal model working"
author: "Tourism Economics"
date: "February 12, 2016"
output: html_document
---


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(arlodr)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(readr, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
library(seasonal, warn.conflicts=FALSE)
require(xtable)
require(ggplot2)
require(knitr)
require(RColorBrewer)
require(scales)
library(lubridate)
library(purrr)
library(lazyeval)
library(forecast)


# If I put autodep=TRUE as an argument in the following then I get
# some issues with the graphs, such as grey backgrounds and changes
# in labeling. Maybe it somehow impacts the theme?
#knitr::opts_chunk$set(cache=TRUE, cache.path='output_data/cache/', dpi=96)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, cache.path='../output_data/cache/', autodep=TRUE, fig.path='../output_data/figure_exp_seasonal_short/fig-',  fig.show='asis', fig.keep='high')
# tells it to keep going even if there is an error
knitr::opts_chunk$set(error = TRUE)

```


```{r}

load(file="~/Project/R projects/lodfor/output_data/holiday_regress.Rdata")
```

```{r}
# unpack lists

a <- c(val_list, mem_list, eas_list, jlf_list, augend_list, sepstr_list, hlw_list, vet_list, thk_list, chr_list)

varNames <- names(a)
for (i in seq_along(varNames)) {
print(varNames[i])
 hold_a <- a[[i]]
 # names the vector, this uses the first object as a name for the second object
 assign(paste(varNames[i], sep=""), hold_a)
}

```



```{r echo=FALSE, message=FALSE, warning=FALSE}
load("~/Project/R projects/lodfor/output_data/ushist_m.Rdata")
load("~/Project/R projects/lodfor/output_data/ushist_q.Rdata")
load("~/Project/R projects/lodfor/output_data/ushist_a.Rdata")

```

```{r define_theme, echo=FALSE, message=FALSE, warning=FALSE}
#Displays palette and sets the ts1 theme

#display.brewer.pal(10, "RdBu")
mypallete <- brewer.pal( 10 , "RdBu" )
colors_ts1 <- scale_colour_manual(values = mypallete[c(10, 1, 5, 4)])
# I don't use the following, was an example I saw somewhere maybe as a way to 
# interpolate additional palette colors
#mypal <- colorRampPalette( brewer.pal( 10 , "RdBu" ) )
theme_ts1 <- function(base_size = 12, base_family = "") {
  theme_classic(base_size = base_size) %+replace%
    theme(                                 
      panel.grid.major.x=element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.minor.y=element_blank(),
      axis.ticks.y=element_line(size=.2),
      axis.ticks.x=element_line(size=.2),
      panel.background=element_blank(),
      legend.title=element_blank(),
      legend.key=element_rect(fill="white", colour = "white"),
      legend.key.size=unit(1, "cm"),
      legend.text=element_text(size=rel(1)),
      #legend.position = c(1, .1),
      legend.justification = "right",
      axis.line=element_line(size=.2),
      axis.title.y=element_blank(),
      axis.title.x=element_blank(),
      axis.text=element_text(color="black",size=rel(.7)),
      # following seems to adjust y-axis title
      plot.title=element_text(size=base_size * .8, face="plain",hjust=0,vjust=1)
    #theme(plot.title = element_text(lineheight=.8, face="bold"))
    )
}
theme_set(theme_ts1())
```

```{r}

segrun <- c("upmus_demd")

```


### Set up data to run using: `r segrun`
```{r echo=FALSE, message=FALSE, warning=FALSE}

# shorten dates
#ushist_m_1 <- ushist_m["1987-01-01/2015-09-01"]

# sets up a single series to use for a few examples
demo_df <- ushist_m %>%
  data.frame(date=time(.), .) %>%
  select_("date", segrun)

# sets up a similar annual data frame
demo_df_a <- ushist_a %>%
  data.frame(date=time(.), .) %>%
  mutate(upmus_demd = upmus_demt / 365) %>%
  select_("date", segrun)
  
x <- demo_df_a %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

tempa <- x %>%
  data.frame(date=time(.), .) %>%
  rename_("sa1a" = segrun) %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  as.xts(order.by = as.yearmon(index(.)))

# use annual to create data frame
# monthly series that is the annual value for each month
# steps are a bit lengthy, but this is the way I could figure out to get the 
# annual data turned into monthly, filling all the way through the current year.
tempa_1 <- tempa %>%
  data.frame(date=time(.), .) %>%
  mutate(year = year(date)) %>%
  select(-date)

# creates a sequence of dates
sq <- seq(as.Date(start(tempa)), as.Date(end(tempa), frac = 1), by = "month") %>%
  as.yearmon(.) %>%
  data.frame()
colnames(sq) <- c("date")
sq <- sq %>%
  mutate(year = year(date))

# join the data onto the data
tempa_by_m  <- sq %>%
  left_join(tempa_1, by=c("year")) %>%
  filter(!(is.na(sa1a))) %>%
  select(-year) %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

```
  
  
  
```{r}

# # Full history, create time series
# x <- demo_df %>%
#   read.zoo(regular=TRUE, drop=FALSE) %>%
#   xts() %>%
#   convertIndex("yearmon")
# temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# # with more regressors
# m3 <- seas(x = temp_ser_ts,
#            regression.variables = c("tdnolpyear", "ls2001.Sep"),
#            forecast.maxlead = 30,# extends 30 periods ahead
#            seats.appendfcst="yes",
#            xreg = cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m, 
#                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m), 
#            regression.usertype = c("holiday", "holiday2", "holiday3"),
#            regression.aictest = NULL)
# summary(m3)
# 
# # try it by creating on object to hold the regressors
# hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m, 
#                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m)
# m4 <- seas(x = temp_ser_ts,
#            regression.variables = c("tdnolpyear", "ls2001.Sep"),
#            forecast.maxlead = 30,# extends 30 periods ahead
#            seats.appendfcst="yes",
#            xreg = hold_reg, 
#            regression.usertype = c("holiday", "holiday2", "holiday3"),
#            regression.aictest = NULL)
# summary(m4)
# 
# # try with a different set of regressors
# hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m, 
#                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m)
# m5 <- seas(x = temp_ser_ts,
#            regression.variables = c("tdnolpyear", "ls2001.Sep"),
#            forecast.maxlead = 30,# extends 30 periods ahead
#            seats.appendfcst="yes",
#            xreg = hold_reg, 
#            regression.usertype = c("holiday", "holiday2", "holiday3"),
#            regression.aictest = NULL)
# summary(m5)
# 
# # try another group
# hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m, 
#                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m, hlw_fss_ts_m)
# m6 <- seas(x = temp_ser_ts,
#            regression.variables = c("tdnolpyear", "ls2001.Sep"),
#            forecast.maxlead = 30,# extends 30 periods ahead
#            seats.appendfcst="yes",
#            xreg = hold_reg, 
#            regression.usertype = c("holiday", "holiday2", "holiday3"),
#            regression.aictest = NULL)
# summary(m6)

# try another group
# hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m,
#                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m, hlw_fss_ts_m,
#                   thk_2826_ts_m) #
# m7 <- seas(x = temp_ser_ts,
#            regression.variables = c("tdnolpyear", "ls2001.Sep"),
#            forecast.maxlead = 30,# extends 30 periods ahead
#            seats.appendfcst="yes",
#            xreg = hold_reg,
#            regression.usertype = c("holiday", "holiday2", "holiday3"),
#            regression.aictest = NULL)
# summary(m7)
# 
# # try another group
# hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m,
#                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m, hlw_fss_ts_m,
#                   thk_2826_ts_m, chr_fssm_ts_m) 
# m8 <- seas(x = temp_ser_ts,
#            regression.variables = c("tdnolpyear", "ls2001.Sep"),
#            forecast.maxlead = 30,# extends 30 periods ahead
#            seats.appendfcst="yes",
#            xreg = hold_reg,
#            regression.usertype = c("holiday", "holiday2", "holiday3"),
#            regression.aictest = NULL)
# summary(m8)
# 
# # try another group
# hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m,
#                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m, hlw_fss_ts_m,
#                   thk_2826_ts_m, chr_fri_ts_m) 
# m9 <- seas(x = temp_ser_ts,
#            regression.variables = c("tdnolpyear", "ls2001.Sep"),
#            forecast.maxlead = 30,# extends 30 periods ahead
#            seats.appendfcst="yes",
#            xreg = hold_reg,
#            regression.usertype = c("holiday", "holiday2", "holiday3"),
#            regression.aictest = NULL)
# summary(m9)

```

## Training period through 2012, test on 2013

Estimate seasonal model on a training period, use it to look at a test period
```{r}

# Full history
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

# shorten dates, create time series
x<- x["1987-01-01/2015-12-01"]
#temp_ser_ts <- as.ts(x, start = start(x), end = end(x))


# timeSlices <- createTimeSlices(1:nrow(x), 
#                    initialWindow = 332, horizon = 12, fixedWindow = FALSE)
# 
# df_1 <- data_frame(train_slices = timeSlices[[1]], test_slices = timeSlices[[2]])
# str(df_1)
# str(x)
# train_slices <- x[trainSlices[[1]],]
# head(train_slices)
# 
# 
# df_2 <- df_1 %>%
#   mutate(hold_df)
# 
# economics[trainSlices[[1]],]
################

f_temp1 <- function(x, start, end) {
  .dots <- list(~date <= end)
  filter_(x, .dots = .dots)
}

f_convert_to_ts <- function(x) {
  x <- x %>%
    read.zoo(regular=TRUE, drop=FALSE) %>%
    xts() %>%
    convertIndex(value="yearmon") %>%
    as.ts(start = start(.), end = end(.))
}

f_seas1<- function(x) {
  seas(x,
           transform.function = "log",
           regression.variables = c("tdnolpyear", "ls2001.Sep"),
           forecast.maxlead = 48,# extends 30 periods ahead
           seats.appendfcst="yes",
           xreg = hold_reg, 
           regression.usertype = c("holiday", "holiday2", "holiday3"),
           regression.aictest = NULL)
}

f_hw<- function(x) {
  y <- hw(x, seasonal="multiplicative")
  # as.zoo()%>% 
  #  as.xts(order.by = as.yearmon(index(.)))  %>%
  #  data.frame(date=time(.), .)
}

f_extract_series <- function(x) {
merge(
   # creates and merges invididual zoo series
   og = as.zoo(original(x)),
   sa = as.zoo(final(x)),
   sf = as.zoo(series(x, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.))) %>%
  data.frame(date=time(.), .) %>%
  mutate(date = as.Date(date)) %>%
  # calculate a forecast of non-seasonally adjusted
  mutate(og_fct = sa*sf)
}

f_extract_hw <- function(X) {
 x$mean #%>%
   #as.zoo() 
 
 # %>% 
 #   as.xts(order.by = as.yearmon(index(.))) %>%
 #   data.frame(date=time(.), .) %>%
 #   mutate(date = as.Date(date))
 #   colnames(b)[2] <- c("oghw_fct")
    
}

f_ready_1 <- function(x, start_var, end_var) {
  x <- x %>%
  mutate(train_start = as.Date(start_var),
            train_end = as.Date(end_var)) %>%
  # calculates number of months since end of training period
  mutate(proj_month = as.period(interval(ymd(date), ymd(train_end)), unit="months")) %>%
  mutate(proj_month = as.character(proj_month)) %>%
  mutate(proj_month = gsub("m 0d 0H 0M 0S","",proj_month)) %>%
  mutate(proj_month = gsub("0S",0,proj_month)) %>%
  mutate(proj_month = as.integer(proj_month)*-1)
}

x2 <- x %>%
  data.frame(date=time(.), .) %>%
  mutate(date = as.Date(date))

df_1 <- data_frame(start_1 = as.Date("1987-01-01"), end_1 = seq(from =2010, to = 2014), hold = list(x2,x2,x2,x2,x2))


df_2 <- df_1 %>%
  # formats date
  mutate(end_1 = as.Date(paste0(end_1, "-12-01"))) %>%
  # uses the function I defined to filter based on start and end dates
  mutate(hold_ts = pmap(list(x=hold, start=start_1, end=end_1), f_temp1)) %>% 
  # converts to time series
  mutate(hold_ts = map(hold_ts, f_convert_to_ts))

# temporary look
# a <- df_3$hold_df3[[2]]
# str(a)
# b <- as.ts(a, start = start(a), end = end(a))
# b

# another temporary look
# a <- df_4$hold_ts[1]
# str(a)

# regressors
hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m,
                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m, hlw_fss_ts_m,
                   thk_2826_ts_m, chr_fri_ts_m)  

#estimate models
df_3 <- df_2 %>%
  mutate(model = map(hold_ts, f_seas1)) %>%
  # fit a holt winters seasonal model
  mutate(model_hw = map(hold_ts, f_hw))

ab <-df_3$model_hw[4]
str(ab)
ba <- ab[[1]]


ba  <- ab[[1]][[2]] %>%
  as.zoo()%>% 
   as.xts(order.by = as.yearmon(index(.))) 

ba <- ab$mean %>%
   as.zoo() %>% 
   as.xts(order.by = as.yearmon(index(.))) %>%
   data.frame(date=time(.), .) %>%
   mutate(date = as.Date(date))
   colnames(ba)[2] <- c("oghw_fct")
 
     ba
c <- f_extract_hw(a)

#extract series
df_4 <- df_3 %>%
  mutate(selected_series = map(model, f_extract_series)) %>%
  # modified data frame of selected series a bit
  mutate(selected_series = pmap(list(x=selected_series, start_var=start_1, end_var=end_1), f_ready_1)) %>%
  # from the hw models stored in a column, extract the second element, which is the mean
  mutate(hw_mean = map(model_hw, "mean")) 

# for looking at an individual element
a_1 <- df_4$hw_mean[[1]]
str(a_1)

######
#
# extract the hw predictions
# I couldn't manage to do this as a map step.Keeping getting hung up with
# having a "list of 1" containing a time series.

# this works
a_1 <- df_4$hw_mean
d_1 <- NULL
for (i in 1:5) {
  e_1 <- xts(a_1[[i]], order.by = as.yearmon(index(a_1[[i]])))
  #print(e_1)
  d_1 <- cbind(d_1, e_1)
}

# convert to a data frame
d_2 <- d_1 %>%
   data.frame(date=time(.), .) %>%
   mutate(date = as.Date(date))
# then I would need to do this to help name it
colnames(d_2) <- c("date", as.character(df_4$end_1))
# tidy
hw_mean_results <- d_2 %>%
  gather(train_end, oghw_fct, -date) %>%
  mutate(train_end = as.Date(train_end))

###
#
# extract estimated factors 

est_factors_1 <- df_4 %>%
  select(selected_series) %>%
  unnest() 

#### add on the hw predictions
# made the assumption there is only one test_start
# if the following is not true, then need to think about the code more. Need
# to bring train_start into the dataframe of hw results to make better join.
stopifnot(1 == length(unique(est_factors_1$train_start)))

est_factors_2 <- est_factors_1 %>%
  left_join(hw_mean_results, by=c("train_end", "date"))

####
#
# could I compare to a smoothed version of the actual series

# first rename the series
colnames(x2)[2] <- c("og_act")

x3 <- x2 %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

# fit a STL decomposition
# STL stands for Seasonal and Trend decomposition using Loess
fit_stl <- stl(x3,s.window="periodic")

# # extract the fitted one-step ahead forecast values
# hw <- fit2$fitted %>%
#   xts(order.by=(index(x3)))
# # name the series
# colnames(trend)[1] <- c("oghw_act")

a <- fit_stl$time.series
trend <- a[,"trend"] %>%
  xts(order.by=(index(x3)))

# name the series
colnames(trend)[1] <- c("ogstl_act")
# merge with original
x4 <- x3 %>%
  merge(trend) %>%
  data.frame(date=time(.), .) %>%
  mutate(date = as.Date(date))


# graph it to tak a look
tograph <- x4 %>%
  gather(variable, value, -date)
p1 <- ggplot(data=tograph, aes(x=date, y=value, color=variable))
p1 <- p1 + geom_line()
p1

# join the original data onto the test projections
est_factors_3 <- est_factors_2 %>%
  left_join(x4, by="date") %>%
  arrange(train_end, date) %>%
  group_by(train_end) %>%
  mutate(og_act_pchya = og_act / lag(og_act,12)-1) %>%
  # calculate based on hw forecast
  mutate(oghw_fct_pchya = oghw_fct / lag(og_act,12)-1) %>%
  # calculate a forecast based on the stl trend
  mutate(ogstl_fct = ogstl_act*sf) %>%
  mutate(ogstl_fct_pchya = ogstl_fct / lag(og_act,12)-1) %>%
  # calculate a trend based pchya based on the stl trend
  mutate(ogstl_act_pchya = ogstl_act / lag(ogstl_act,12)-1) %>%
  ungroup()

est_factors_4 <- est_factors_3 %>%
    filter(proj_month >= 1 & proj_month <= 12)

# looking at levels
tograph <- est_factors_4 %>%
#  filter(train_end == as.Date("2012-12-01")) %>%
  gather(variable, value, -date, -train_start, -train_end, -proj_month) %>%
  filter(variable %in% c("og_act", "ogstl_act", "ogstl_fct"))

p1 <- ggplot(data=tograph, aes(x=proj_month, y=value, color=variable))
p1 <- p1 + geom_line()
p1 <- p1 + facet_wrap(~train_end)
p1

# looking at pchya
tograph <- est_factors_4 %>%
#  filter(train_end == as.Date("2012-12-01")) %>%
  gather(variable, value, -date, -train_start, -train_end, -proj_month) %>%
  filter(variable %in% c("og_act_pchya", "ogstl_act_pchya", "ogstl_fct_pchya"))

p1 <- ggplot(data=tograph, aes(x=proj_month, y=value, color=variable))
p1 <- p1 + geom_line()
p1 <- p1 + facet_wrap(~train_end)
p1


# wow, lot's of progress up to this point!





# mmod <- seas(x = temp_ser_ts,
#            transform.function = "log",
#            regression.variables = c("tdnolpyear", "ls2001.Sep"),
#            forecast.maxlead = 48,# extends 30 periods ahead
#            seats.appendfcst="yes",
#            xreg = hold_reg, 
#            regression.usertype = c("holiday", "holiday2", "holiday3"),
#            regression.aictest = NULL)
summary(mmod)

# use monthly to create data frame
tempm <- merge(
   # creates and merges invididual zoo series
   oa1 = as.zoo(original(mmod)),
   sa1 = as.zoo(final(mmod)),
   sa1_sf = as.zoo(series(mmod, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.)))

# replace original series with the full original series
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")
colnames(x) <- c("oa1")
tempm <- tempm %>%
  data.frame(date=time(.), .) %>%
  select(-oa1) %>%
    read.zoo(regular=TRUE, drop=FALSE) %>%
  as.xts(order.by = as.yearmon(index(.))) %>%
  merge(., x)

# merge together annual 
tempm1 <- merge(tempm, tempa_by_m)

# create a monthly series that is calculated as the
# sa-adjusted quarterly data times the monthly seasonal
# factors
tempm1 <- tempm1 %>%
  transform(oa1a = sa1a * sa1_sf) %>%
  transform(oa1_pchya = oa1 / lag(oa1, 12) - 1) %>%
  transform(oa1a_pchya = oa1a / lag(oa1a, 12) - 1) %>%
  transform(oa1a_pchya_demi = oa1a / lag(oa1, 12) - 1) %>%
  xts()

# look at data
a <- tempm1 %>%
  data.frame(date=time(.), .) %>%
  filter(date <= "2013-12-01") %>%
  mutate_each(funs(sprintf("%1.2f%%",100*.)), contains("_pchya"))
kable(tail(a, n=12), digits = 3, row.names=FALSE, align = c("l", rep("r", 7)))

```

### Monthly data, comparison between original and oa1a
Percentage change from prior year  
```{r}
# look at it on a percentage change from prior year basis
temp_graph <- tempm1 %>%
  data.frame(date=as.Date(time(.)), .) %>%
  #mutate_each(funs(pchya = . / lag(., 12) - 1), -date) %>%
  gather("variable", "value", -date) %>%
  filter(variable %in% c("oa1_pchya", "oa1a_pchya_demi")) %>%
  filter(date <= "2013-12-01" & date > "2011-01-01")

p1 <- temp_graph %>%
   ggplot()
p1 <- p1 + geom_line(data = temp_graph, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40", mypallete[7:7]))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 + scale_y_continuous(labels=percent)
p1 <- p1 + geom_vline(xintercept=as.numeric(temp_graph$date[23]), linetype=4)
p1

```

## Training period through 2013, test on 2014

Estimate seasonal model on a training period, use it to look at a test period
```{r}

# Full history
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

# shorten dates, create time series
x<- x["1987-01-01/2013-12-01"]
temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# estimate model
hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m,
                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m, hlw_fss_ts_m,
                   thk_2826_ts_m, chr_fri_ts_m)  

mmod <- seas(x = temp_ser_ts,
             transform.function = "log",
           regression.variables = c("tdnolpyear", "ls2001.Sep"),
           forecast.maxlead = 48,# extends 30 periods ahead
           seats.appendfcst="yes",
         #  xreg = hold_reg, 
          # regression.usertype = c("holiday", "holiday2", "holiday3"),
           regression.aictest = NULL)
summary(mmod)

# use monthly to create data frame
tempm <- merge(
   # creates and merges invididual zoo series
   oa1 = as.zoo(original(mmod)),
   sa1 = as.zoo(final(mmod)),
   sa1_sf = as.zoo(series(mmod, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.)))

#as.zoo(final(mmod))
# p1 <- tempm %>%
#     data.frame(date=time(.), .) %>%
#   mutate(date = as.Date(date)) %>%
#   filter(date <= as.Date("2013-12-01") & date >= as.Date("2010-01-01")) %>%
#  ggplot(data = ., aes(x = date, y = sa1))
# p1 <- p1 + geom_line() 
# p1

# replace original series with the full original series
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")
colnames(x) <- c("oa1")
tempm <- tempm %>%
  data.frame(date=time(.), .) %>%
  select(-oa1) %>%
    read.zoo(regular=TRUE, drop=FALSE) %>%
  as.xts(order.by = as.yearmon(index(.))) %>%
  merge(., x)

# merge together annual 
tempm1 <- merge(tempm, tempa_by_m)

# create a monthly series that is calculated as the
# sa-adjusted quarterly data times the monthly seasonal
# factors
tempm1 <- tempm1 %>%
  transform(oa1a = sa1a * sa1_sf) %>%
  transform(oa1_pchya = oa1 / lag(oa1, 12) - 1) %>%
  transform(oa1a_pchya = oa1a / lag(oa1a, 12) - 1) %>%
  transform(oa1a_pchya_demi = oa1a / lag(oa1, 12) - 1) %>%
  xts()

# look at data
a <- tempm1 %>%
  data.frame(date=time(.), .) %>%
  filter(date <= "2014-12-01") %>%
  mutate_each(funs(sprintf("%1.2f%%",100*.)), contains("_pchya"))
kable(tail(a, n=12), digits = 3, row.names=FALSE, align = c("l", rep("r", 7)))

```

### Monthly data, comparison between original and oa1a
Percentage change from prior year  
```{r}
# look at it on a percentage change from prior year basis
temp_graph <- tempm1 %>%
  data.frame(date=as.Date(time(.)), .) %>%
  #mutate_each(funs(pchya = . / lag(., 12) - 1), -date) %>%
  gather("variable", "value", -date) %>%
  filter(variable %in% c("oa1_pchya", "oa1a_pchya_demi")) %>%
  filter(date <= "2014-12-01" & date > "2012-01-01")

p1 <- temp_graph %>%
   ggplot()
p1 <- p1 + geom_line(data = temp_graph, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40", mypallete[7:7]))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 +  scale_y_continuous(labels=percent)
p1 <- p1 + geom_vline(xintercept=as.numeric(temp_graph$date[23]), linetype=4)
p1

```


## Training period through 2014, test on 2015

Estimate seasonal model on a training period, use it to look at a test period
```{r}

# Full history
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

# shorten dates, create time series
x<- x["1987-01-01/2014-12-01"]
temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# estimate model
hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m,
                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m, hlw_fss_ts_m,
                   thk_2826_ts_m, chr_fri_ts_m)  

mmod <- seas(x = temp_ser_ts,
             transform.function = "log",
           regression.variables = c("tdnolpyear", "ls2001.Sep"),
           forecast.maxlead = 48,# extends 30 periods ahead
           seats.appendfcst="yes",
           xreg = hold_reg, 
           regression.usertype = c("holiday", "holiday2", "holiday3"),
           regression.aictest = NULL)
summary(mmod)

# use monthly to create data frame
tempm <- merge(
   # creates and merges invididual zoo series
   oa1 = as.zoo(original(mmod)),
   sa1 = as.zoo(final(mmod)),
   sa1_sf = as.zoo(series(mmod, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.)))

#as.zoo(final(mmod))
# p1 <- tempm %>%
#     data.frame(date=time(.), .) %>%
#   mutate(date = as.Date(date)) %>%
#   filter(date <= as.Date("2013-12-01") & date >= as.Date("2010-01-01")) %>%
#  ggplot(data = ., aes(x = date, y = sa1))
# p1 <- p1 + geom_line() 
# p1

# replace original series with the full original series
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")
colnames(x) <- c("oa1")
tempm <- tempm %>%
  data.frame(date=time(.), .) %>%
  select(-oa1) %>%
    read.zoo(regular=TRUE, drop=FALSE) %>%
  as.xts(order.by = as.yearmon(index(.))) %>%
  merge(., x)

# merge together annual 
tempm1 <- merge(tempm, tempa_by_m)

# create a monthly series that is calculated as the
# sa-adjusted quarterly data times the monthly seasonal
# factors
tempm1 <- tempm1 %>%
  transform(oa1a = sa1a * sa1_sf) %>%
  transform(oa1_pchya = oa1 / lag(oa1, 12) - 1) %>%
  transform(oa1a_pchya = oa1a / lag(oa1a, 12) - 1) %>%
  transform(oa1a_pchya_demi = oa1a / lag(oa1, 12) - 1) %>%
  xts()

# look at data
a <- tempm1 %>%
  data.frame(date=time(.), .) %>%
  filter(date <= "2015-12-01") %>%
  mutate_each(funs(sprintf("%1.2f%%",100*.)), contains("_pchya"))
kable(tail(a, n=12), digits = 3, row.names=FALSE, align = c("l", rep("r", 7)))

```

### Monthly data, comparison between original and oa1a
Percentage change from prior year  
```{r}
# look at it on a percentage change from prior year basis
temp_graph <- tempm1 %>%
  data.frame(date=as.Date(time(.)), .) %>%
  #mutate_each(funs(pchya = . / lag(., 12) - 1), -date) %>%
  gather("variable", "value", -date) %>%
  filter(variable %in% c("oa1_pchya", "oa1a_pchya_demi")) %>%
  filter(date <= "2015-12-01" & date > "2013-01-01")

p1 <- temp_graph %>%
   ggplot()
p1 <- p1 + geom_line(data = temp_graph, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40", mypallete[7:7]))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 +  scale_y_continuous(labels=percent)
p1 <- p1 + geom_vline(xintercept=as.numeric(temp_graph$date[23]), linetype=4)
p1

```

## Training period through 2015, test on 2016


Estimate seasonal model on a training period, use it to look at a test period
```{r}

# Full history
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")

# shorten dates, create time series
x<- x["1987-01-01/2015-12-01"]
temp_ser_ts <- as.ts(x, start = start(x), end = end(x))

# estimate model
hold_reg <- cbind(eas_7_1_ts_m, val_fs_ts_m, augend_ss_ts_m,
                        sepstr_ss_ts_m, sepstr_mon_ts_m, mempost_m, jlf_fssm_ts_m, hlw_fss_ts_m,
                   thk_2826_ts_m, chr_fri_ts_m)  

mmod <- seas(x = temp_ser_ts,
             transform.function = "log",
           regression.variables = c("tdnolpyear", "ls2001.Sep"),
           forecast.maxlead = 48,# extends 30 periods ahead
           seats.appendfcst="yes",
           xreg = hold_reg, 
           regression.usertype = c("holiday", "holiday2", "holiday3"),
           regression.aictest = NULL)
summary(mmod)

# use monthly to create data frame
tempm <- merge(
   # creates and merges invididual zoo series
   oa1 = as.zoo(original(mmod)),
   sa1 = as.zoo(final(mmod)),
   sa1_sf = as.zoo(series(mmod, "s16"))) %>%
   # converts to xts
   as.xts(order.by = as.yearmon(index(.)))

#as.zoo(final(mmod))
# p1 <- tempm %>%
#     data.frame(date=time(.), .) %>%
#   mutate(date = as.Date(date)) %>%
#   filter(date <= as.Date("2013-12-01") & date >= as.Date("2010-01-01")) %>%
#  ggplot(data = ., aes(x = date, y = sa1))
# p1 <- p1 + geom_line() 
# p1

# replace original series with the full original series
x <- demo_df %>%
  read.zoo(regular=TRUE, drop=FALSE) %>%
  xts() %>%
  convertIndex("yearmon")
colnames(x) <- c("oa1")
tempm <- tempm %>%
  data.frame(date=time(.), .) %>%
  select(-oa1) %>%
    read.zoo(regular=TRUE, drop=FALSE) %>%
  as.xts(order.by = as.yearmon(index(.))) %>%
  merge(., x)

# merge together annual 
tempm1 <- merge(tempm, tempa_by_m)

# create a monthly series that is calculated as the
# sa-adjusted quarterly data times the monthly seasonal
# factors
tempm1 <- tempm1 %>%
  transform(oa1a = sa1a * sa1_sf) %>%
  transform(oa1_pchya = oa1 / lag(oa1, 12) - 1) %>%
  transform(oa1a_pchya = oa1a / lag(oa1a, 12) - 1) %>%
  transform(oa1a_pchya_demi = oa1a / lag(oa1, 12) - 1) %>%
  xts()

# look at data
a <- tempm1 %>%
  data.frame(date=time(.), .) %>%
  filter(date <= "2016-12-01") %>%
  mutate_each(funs(sprintf("%1.2f%%",100*.)), contains("_pchya"))
kable(tail(a, n=12), digits = 3, row.names=FALSE, align = c("l", rep("r", 7)))

```

### Monthly data, comparison between original and oa1a
Percentage change from prior year  
```{r}
# look at it on a percentage change from prior year basis
temp_graph <- tempm1 %>%
  data.frame(date=as.Date(time(.)), .) %>%
  #mutate_each(funs(pchya = . / lag(., 12) - 1), -date) %>%
  gather("variable", "value", -date) %>%
  filter(variable %in% c("oa1_pchya", "oa1a_pchya_demi")) %>%
  filter(date <= "2016-12-01" & date > "2014-01-01")

p1 <- temp_graph %>%
   ggplot()
p1 <- p1 + geom_line(data = temp_graph, aes(x = date, y = value, color=variable), size = .7) 
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey40", mypallete[7:7]))
p1 <- p1 + labs(colour = "", x = "")
p1 <- p1 + ggtitle("Title")
p1 <- p1 + theme(legend.position = "bottom")
p1 <- p1 +  scale_y_continuous(labels=percent)
p1 <- p1 + geom_vline(xintercept=as.numeric(temp_graph$date[23]), linetype=4)
p1

```





