---
title: "Seasonal model working graphs"
author: "Tourism Economics"
date: "March 30, 2016"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(arlodr)
library(xts, warn.conflicts=FALSE)
library(dplyr, warn.conflicts=FALSE)
library(readr, warn.conflicts=FALSE)
library(tidyr, warn.conflicts=FALSE)
library(seasonal, warn.conflicts=FALSE)
require(xtable)
require(ggplot2)
require(knitr)
require(RColorBrewer)
require(scales)
library(lubridate)
library(purrr)
library(lazyeval)
library(forecast)


require(directlabels)
require(gridExtra)
require(gtable)
require(Cairo)
require(grid)

# If I put autodep=TRUE as an argument in the following then I get
# some issues with the graphs, such as grey backgrounds and changes
# in labeling. Maybe it somehow impacts the theme?

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, cache.path="/output_data/cache/", autodep=TRUE, fig.path="/output_data/figure_sea_model_work_structure/fig-",  fig.show="asis", fig.keep="high")

# tells it to keep going even if there is an error
knitr::opts_chunk$set(error = TRUE)

# increases resolution of figures being shown in html
opts_chunk$set(fig.width = 9, fig.height = 5, dpi = 144)
```


```{r set segrun}

segrun <- c("totus_adr")

```
Series being run is `r segrun`.

```{r}


fpath <- c("~/Project/R projects/lodfor/") 

load(file=paste0(fpath,"output_data/out_seamod_", segrun, ".Rdata"))
work_seamod <- out_seamod

```



```{r define_theme, echo=FALSE, message=FALSE, warning=FALSE}
#Displays palette and sets the ts1 theme

#display.brewer.pal(10, "RdBu")
mypallete <- brewer.pal( 10 , "RdBu" )
colors_ts1 <- scale_colour_manual(values = mypallete[c(10, 1, 5, 4)])
# I don't use the following, was an example I saw somewhere maybe as a way to 
# interpolate additional palette colors
#mypal <- colorRampPalette( brewer.pal( 10 , "RdBu" ) )
theme_ts1 <- function(base_size = 12, base_family = "") {
  theme_classic(base_size = base_size) %+replace%
    theme(                                 
      panel.grid.major.x=element_blank(),
      panel.grid.minor.x=element_blank(),
      panel.grid.minor.y=element_blank(),
      axis.ticks.y=element_line(size=.2),
      axis.ticks.x=element_line(size=.2),
      panel.background=element_blank(),
      legend.title=element_blank(),
      legend.key=element_rect(fill="white", colour = "white"),
      legend.key.size=unit(1, "cm"),
      legend.text=element_text(size=rel(1)),
      #legend.position = c(1, .1),
      legend.justification = "right",
      axis.line=element_line(size=.2),
      axis.title.y=element_blank(),
      axis.title.x=element_blank(),
      axis.text=element_text(color="black",size=rel(.7)),
      # following seems to adjust y-axis title
      plot.title=element_text(size=base_size * .8, face="plain",hjust=0,vjust=1)
    #theme(plot.title = element_text(lineheight=.8, face="bold"))
    )
}
theme_set(theme_ts1())
```


```{r}

out_hold <- work_seamod %>%
  unnest(factorsout) %>%
  mutate(ogstl_fct_pchya_absd = abs(ogstl_fct_pchya - og_act_pchya)) %>%
  mutate(ogstl_act_pchya_absd = abs(ogstl_act_pchya - og_act_pchya)) 

out_hold_2 <- out_hold %>%
  # filter to the first year of projections
  filter(proj_month >= 1 & proj_month <= 12) 

```


```{r}

temp_a <- out_hold %>%
  filter(model_id == "model_8" ) %>%
  filter(train_end == as.Date("2014-12-01")) %>%
  filter(date >= as.Date("2003-01-01")) %>%
  filter(date <= as.Date("2014-12-01"))
  
tograph <- temp_a %>%
  select(-model_id, -train_start, -train_end) %>%
  gather(variable, value, -date, -proj_month) %>%
  mutate(value = as.numeric(value)) %>%
  filter(variable %in% c("og_act", "ogstl_act", "ogstl_fct")) 

# all years
p1 <- ggplot(data=tograph, aes(x=date, y=value, group=variable, size=variable, color=variable))
p1 <- p1 + geom_line()
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey70", mypallete[9:9]))
p1 <- p1 + scale_size_manual(values=c(.7, 1.2, .7))
p1 <- p1 + scale_alpha_manual(values=c(1, .5, 1))
# adds points for one group
p1 <- p1 + geom_point(data = subset(tograph, variable %in% c("og_act")), size=1, color=mypallete[3:3])
# points in blank points for all, so that we can override the legend
p1 <- p1 + geom_point(size=0)
# overrides legend to also show shape for the first series
p1 <- p1 + guides(color=guide_legend(override.aes=list(shape=c(16,NA,NA))))

# shorter period
tograph_1 <- tograph %>%
  filter(date >= as.Date("2012-01-01"))
p2 <- ggplot(data=tograph_1, aes(x=date, y=value, group=variable, size=variable, color=variable))
p2 <- p2 + geom_line()
p2 <- p2 + scale_color_manual(values = c(mypallete[3:3], "grey70", mypallete[9:9]))
p2 <- p2 + scale_size_manual(values=c(.7, 1.2, .7))
p2 <- p2 + scale_alpha_manual(values=c(1, .5, 1))
# adds points for one group
p2 <- p2 + geom_point(data = subset(tograph_1, variable %in% c("og_act")), size=1, color=mypallete[3:3])
# points in blank points for all, so that we can override the legend
p2 <- p2 + geom_point(size=0)
# overrides legend to also show shape for the first series
p2 <- p2 + guides(color=guide_legend(override.aes=list(shape=c(16,NA,NA))))


tograph <- temp_a %>%
  select(-model_id, -train_start, -train_end) %>%
  gather(variable, value, -date, -proj_month) %>%
  mutate(value = as.numeric(value)) %>%
  filter(variable %in% c("sa", "ogstl_act")) 

p3 <- ggplot(data=tograph, aes(x=date, y=value, group=variable, size=variable, color=variable))
p3 <- p3 + geom_line()
p3 <- p3 + scale_color_manual(values = c("grey70", mypallete[9:9]))
p3 <- p3 + scale_size_manual(values=c(1.2, .8))
p3 <- p3 + scale_alpha_manual(values=c(.5, 1))
p3 <- p3 + scale_y_log10()

```

### Comparison over time
* og_act: actual results (non-seasonally adjusted)  
* ogstl_act: actual results, but smoothed using a seasonal decomposition by loess (stl)  
* ogstl_fct: forecast, based on the ogstl_act series multiplied by the estimated seasonal factors  
```{r}
p1
```


### Comparison over shorter period of time
```{r}
p2
```


### Comparison of smoothed series
```{r}
p3
```


```{r}

selected <- out_hold_2 %>%
  mutate(year = year(date)) %>%
  mutate(month = month(date, label=TRUE, abbr= FALSE)) %>%
  mutate(month = factor(month, month.name)) %>%
  mutate(month_num = month(date, label=FALSE)) %>%
  filter(model_id == "model_8")

```

### Graphs of levels by year
```{r}

######
#
# graphing the results

# looking at levels
tograph <- selected %>%
  gather(variable, value, -date, -year, -month,  -month_num, -proj_month, -model_id) %>%
  mutate(value = as.numeric(value)) %>%
  filter(variable %in% c("og_act", "ogstl_act", "ogstl_fct")) 

# function to use in creating a short series of plots

plot1 <- function(df) {
p1 <- ggplot(data=df, aes(x=month_num, y=value, group=variable, size=variable, color=variable))
p1 <- p1 + geom_line()
p1 <- p1 + facet_wrap(~year)
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey70", mypallete[9:9]))
p1 <- p1 + scale_size_manual(values=c(.7, 1.2, .7))
p1 <- p1 + scale_alpha_manual(values=c(1, .5, 1))
# adds points for one group
p1 <- p1 + geom_point(data = subset(df, variable %in% c("og_act")), size=1, color=mypallete[3:3])
# points in blank points for all, so that we can override the legend
p1 <- p1 + geom_point(size=0)
# overrides legend to also show shape for the first series
p1 <- p1 + guides(color=guide_legend(override.aes=list(shape=c(16,NA,NA))))
p1
}

tograph_1 <- tograph %>%
  filter(year >= 2005 & year <= 2007)
plot1(tograph_1)

tograph_1 <- tograph %>%
  filter(year >= 2008 & year <= 2010)
plot1(tograph_1)

tograph_1 <- tograph %>%
  filter(year >= 2011 & year <= 2013)
plot1(tograph_1)

tograph_1 <- tograph %>%
  filter(year >= 2014 & year <= 2016)
plot1(tograph_1)

```

******  
### Graph of percent change from prior year, by year
```{r}

# looking at pchya
tograph <- selected %>%
  gather(variable, value, -date, -year, -month, -month_num, -proj_month, -model_id) %>%
  mutate(value = as.numeric(value)) %>%
  filter(variable %in% c("og_act_pchya", "ogstl_act_pchya", "ogstl_fct_pchya")) # %>%
  #filter(date >= as.Date("2014-01-01") & date <= as.Date("2015-12-01"))
  
tograph_hold <- tograph %>%
    filter(year == 2015)

plot2 <- function(df) {
  # p1 <- ggplot(data=tograph_hold, aes(x=date, y=value, color=variable))
  # p1 <- p1 + geom_line()
  # p1 <- p1 + ggtitle(i)
  # p1 <- p1 + scale_x_date(date_breaks = "1 month", date_labels = "%b")
  # print(p1)
  
p1 <- ggplot(data=df, aes(x=date, y=value, group=variable, size=variable, color=variable))
p1 <- p1 + geom_line()
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey70", mypallete[9:9]))
p1 <- p1 + scale_size_manual(values=c(.7, 1.2, .7))
p1 <- p1 + scale_alpha_manual(values=c(1, .5, 1))
# adds points for one group
p1 <- p1 + geom_point(data = subset(df, variable %in% c("og_act")), size=1, color=mypallete[3:3])
# points in blank points for all, so that we can override the legend
p1 <- p1 + geom_point(size=0)
# overrides legend to also show shape for the first series
p1 <- p1 + guides(color=guide_legend(override.aes=list(shape=c(16,NA,NA))))
p1 <- p1 + scale_x_date(date_breaks = "1 month", date_labels = "%b")
p1 <- p1 + ggtitle(i)
p1 <- p1 + theme(plot.title = element_text(face="bold", size=12, hjust=0))
print(p1)
  
}

years <- unique(tograph$year)
for (i in years) {
  tograph_hold <- tograph %>%
    filter(year == i)
  
plot2(tograph_hold)
  
}
  
  

```

****** 
### Multiple years
Estimates for each year is based on the training period through the prior December. 
```{r}

tograph_2 <- tograph %>%
  filter(date >= as.Date("2014-01-01")) %>% 
  filter(date <= as.Date("2015-12-01")) 

# nothing special, based on approach for some earlier plots above
p1 <- ggplot(data=tograph_2, aes(x=date, y=value, group=variable, size=variable, color=variable))
p1 <- p1 + geom_line()
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey70", mypallete[9:9]))
p1 <- p1 + scale_size_manual(values=c(.7, 1.2, .7))
p1 <- p1 + scale_alpha_manual(values=c(1, .5, 1))
# adds points for one group
p1 <- p1 + geom_point(data = subset(tograph_2, variable %in% c("og_act")), size=1, color=mypallete[3:3])
# points in blank points for all, so that we can override the legend
p1 <- p1 + geom_point(size=0)
# overrides legend to also show shape for the first series
p1 <- p1 + guides(color=guide_legend(override.aes=list(shape=c(16,NA,NA))))
p1 <- p1 + theme(plot.title = element_text(face="bold", size=12, hjust=0))
p1

```

****** 
### Performance by model
Average error of the seasonal factor approach resulting from each model. Each model has a different set of seasonal dummy regressors.  
model_1 only contains the trading day seasonal regressor. The average error is assessed during the 12 months of each test period.  
  
* fct: Mean absolute error (MAE) of the estimates based on seasonal factors (ogstl_fct_pchya) during each test period  
* act: MAE of the smoothed series (ogstl_act_pchya)  
* ratio: fct/act  

In these examples the MAE is equivalent to the average error in percentage points. In other words, the series being evaluated are in terms of the percentage change from the prior year. So if the MAE is 0.01, then it indicates that the estimate is off by an average of one percentage point (e.g. actual change was 3.0% from prior year, estimated change was 2.0% from prior year, implying an error of 1.0 percentage point).  
```{r}

model_sum <- out_hold_2 %>%
  group_by(model_id) %>%
  summarize(fct = mean(ogstl_fct_pchya_absd), act = mean(ogstl_act_pchya_absd)) %>%
  mutate(ratio = fct/act)

kable(model_sum, digits = 4, row.names=FALSE)

```


### Performance of all models averaged by year
Same measures as above, with each MAPE averaged across the 12 months of each year.  
```{r}

model_sum <- out_hold_2 %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(fct = mean(ogstl_fct_pchya_absd), act = mean(ogstl_act_pchya_absd)) %>%
  mutate(ratio = fct/act)

kable(model_sum, digits = 4, row.names=FALSE)

```


### Performance of selected model averaged by month
Same measures as above, with each MAPE averaged for a particular month across all test periods.  
```{r}


model_sum <- selected %>%
  group_by(month) %>%
  summarize(fct = mean(ogstl_fct_pchya_absd), act = mean(ogstl_act_pchya_absd)) %>%
  mutate(ratio = fct/act)

kable(model_sum, digits = 4, row.names=FALSE)

```

****** 
### Evolution of estimates
Compares the:  
* estimates based on a training period through the end of 2013  
* estimates based on the training period through the end of 2014 (which includes actual data)  
* actual data  
This shows that how the estimates of the seasonal factors for a particular year tend to change once data for that year is being used.
```{r}

first_train_end <- as.Date("2013-12-01") 
second_train_end <- as.Date("2014-12-01")

view_ser <- out_hold %>%
  filter(model_id == "model_8") %>%
  filter(date >= as.Date("2014-01-01") & date <= as.Date("2014-12-01")) %>%
  filter(train_end == first_train_end | train_end == second_train_end)


p1 <- ggplot()
p1 <- p1 + geom_line(data = subset(view_ser, train_end == first_train_end), aes(x=date, y=og_act_pchya, color="og_act_pchya"))

p1 <- p1 + geom_line(data = subset(view_ser, train_end == first_train_end), 
                                   aes(x=date, y=ogstl_fct_pchya, color=paste0("ogstl_fct_pchya ", first_train_end)))
p1 <- p1 + geom_line(data = subset(view_ser, train_end == second_train_end), 
                                   aes(x=date, y=ogstl_fct_pchya, color=paste0("ogstl_fct_pchya ", second_train_end)))
p1 <- p1 + scale_color_manual(values = c(mypallete[3:3], "grey70", mypallete[9:9]))
p1 <- p1 + scale_size_manual(values=c(.7, .7, .7))
# adds points for one set of data
p1 <- p1 + geom_point(data = subset(view_ser, train_end == first_train_end), aes(x=date, y=og_act_pchya, color="og_act_pchya"), size=1)
# points in blank points for all, so that we can override the legend
p1 <- p1 + geom_point(size=0)
# overrides legend to also show shape for the first series
p1 <- p1 + guides(color=guide_legend(override.aes=list(shape=c(16,NA,NA))))
p1

# %>%
#   select(-model_id, -train_start) %>%
#   gather(variable, value, -date, -train_end)

# p1 <- ggplot(data=tograph)
# p1 <- p1 + geom_line(aes(x=date, y=sa, color=factor(train_end)))
# p1 <- p1 + geom_line(aes(x=date, y=ogstl_act, color=factor(train_end)))
# p1


```

