---
title: "Untitled"
author: "Tourism Economics"
date: "March 16, 2016"
output: html_document
---

Step 0: Setting up the data and trainControl:(from your question)


```{r}
library(caret)
library(ggplot2)
library(pls)

data(economics)
str(economics)
```

Step 1: Creating the timeSlices for the index of the data:
```{r}
timeSlices <- createTimeSlices(1:nrow(economics), 
                   initialWindow = 36, horizon = 12, fixedWindow = FALSE)
str(timeSlices)
```


This creates a list of training and testing timeSlices.
```{r}
str(timeSlices,max.level = 1)
```

For ease of understanding, I am saving them in separate variable:
```{r}
trainSlices <- timeSlices[[1]]
testSlices <- timeSlices[[2]]

str(trainSlices)
economics[trainSlices[[1]],]
economics[trainSlices[[2]],]

```

Step 2: Training on the first of the trainSlices:
```{r}
plsFitTime <- train(unemploy ~ pce + pop + psavert,
                    data = economics[trainSlices[[1]],],
                    method = "pls",
                    preProc = c("center", "scale"))
```

Step 3: Testing on the first of the trainSlices:
```{r}
pred <- predict(plsFitTime,economics[testSlices[[1]],])
```


Step 4: Plotting:
```{r}
true <- economics$unemploy[testSlices[[1]]]

plot(true, col = "red", ylab = "true (red) , pred (blue)", ylim = range(c(pred,true)))
points(pred, col = "blue") 
```


You can then do this for all the slices:
```{r}
for(i in 1:length(trainSlices)){
  plsFitTime <- train(unemploy ~ pce + pop + psavert,
                      data = economics[trainSlices[[i]],],
                      method = "pls",
                      preProc = c("center", "scale"))
  pred <- predict(plsFitTime,economics[testSlices[[i]],])


  true <- economics$unemploy[testSlices[[i]]]
  plot(true, col = "red", ylab = "true (red) , pred (blue)", 
            main = i, ylim = range(c(pred,true)))
  points(pred, col = "blue") 
}
```


As mentioned earlier, this sort of timeSlicing is done by your original function in one step:

```{r}
 myTimeControl <- trainControl(method = "timeslice",
                               initialWindow = 36,
                               horizon = 12,
                               fixedWindow = TRUE)
 
 plsFitTime <- train(unemploy ~ pce + pop + psavert,
                     data = economics,
                     method = "pls",
                    preProc = c("center", "scale"),
                     trControl = myTimeControl)
 plsFitTime
```

